### 2020-07-28 article.tex
Changed the formating, made some minor edits

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
%
\documentclass[
  12pt,
  a4paper]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={A closer look at fixed effects regression in structural equation modeling using},
  pdfauthor={Henrik Kenneth Andersen},
  colorlinks=true,
  linkcolor=Maroon,
  filecolor=Maroon,
  citecolor=Blue,
  urlcolor=blue,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% ----- Packages

\usepackage{booktabs}
\usepackage{titling}
\usepackage{bm}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{subfig}
\usepackage{tikz}
\usetikzlibrary{arrows}
\usepackage[nofiglist]{endfloat}
%\floatsetup[figure]{capposition=top}
%\usepackage{kbordermatrix}
\usepackage{blkarray}
\usepackage{setspace}
\usepackage{etoolbox}
\doublespacing

% ----- Math operators

\DeclareMathOperator{\E}{\mathbb{E}}
\DeclareMathOperator{\Var}{\mathrm{Var}}
\DeclareMathOperator{\Cov}{\mathrm{Cov}}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

% ----- Graphics

%\graphicspath{{./images/}}

% ----- Equation labels

\mathtoolsset{showonlyrefs}

% ----- Page numbering

\pagenumbering{gobble}

% ----- Spacing

%\renewenvironment{knitrout}{\vspace{10em}}{}

\BeforeBeginEnvironment{equation}{\begin{singlespace}\vspace*{-\baselineskip}}
\AfterEndEnvironment{equation}{\end{singlespace}\noindent\ignorespaces}
\BeforeBeginEnvironment{align}{\begin{singlespace}\vspace*{-\baselineskip}}
\AfterEndEnvironment{align}{\end{singlespace}\noindent\ignorespaces}

\title{\singlespacing A closer look at fixed effects regression in structural
equation modeling using \texttt{lavaan}}
\author{Henrik Kenneth Andersen}
\date{2020-07-27}

\begin{document}
\maketitle
\begin{abstract}
This article provides an in-depth look at fixed effects regression in
the structural equation modeling (SEM) framework, specifically the
application of fixed effects in the \texttt{lavaan} package for
\texttt{R}. It is meant as a applied guide for researchers, covering the
underlying model specification, syntax, and summary output. Online
supplementary materials further discuss various common extentsions to
the basic fixed-effect model, demonstrating how to relax model
assumptions, deal with measurement error in both the dependent and
independent variables, and include time-invariant predictors in a type
of hybrid fixed-/ random effects model. extbf\{Keywords: Fixed effects,
structural equation modeling, lavaan, R, panel analysis\}
\end{abstract}

\newpage

\pagenumbering{arabic} 
\setcounter{page}{1}

\hypertarget{intro}{%
\section{Introduction}\label{intro}}

Several years ago, Curran and Bauer
(\protect\hyperlink{ref-Curran2011}{2011}) reflected positively on the
growing use of panel studies in empirical social research. Some of the
strengths of panel data are well-known, e.g., the ability to establish
temporal precedence, increased statistical power and the reduction of
potential alternative models. However, perhaps the greatest strength of
panel data is that they allow for a more rigorous testing of substantive
theories. Panel data, i.e., repeated measures of the same observed units
(people, schools, firms, countries, etc.), allow researchers to
decompose the error term into a part that stays constant within units
and the part that changes over time. The part that does not change over
time can be seen as the combined effect of all time-invariant influences
(e.g., sex, date of birth, nationality) on the dependent variable. Fixed
effects (FE) regression involves controlling for these time-invariant
influences via a number of various methods. It thus accounts for a
likely and common source of bias.

Structural equation modeling (SEM) is a popular regression framework.
One of its main strengths is its flexibility. Not only can complex
causal structures with multiple dependent variables be tested
simultaneously, but in longitudinal (and, more generally, hierarchical)
studies both time-varying and invariant predictors can be included, and
effects can easily be allowed to vary over time. Thus researchers can
allow for and study effects that increase or fade over time, or that
appear only in specific periods. Beyond that, with the use of latent
variables, SEM provides a way to deal with measurement error and get
closer to the true underlying constructs of interest.

There are a number of articles describing basic concept of panel model
regression, and FE regression in SEM (e.g., Allison
\protect\hyperlink{ref-Allison2011}{2011}; Bollen and Brand
\protect\hyperlink{ref-Bollen2010}{2010}; Teachman et al.
\protect\hyperlink{ref-Teachman2001}{2001}). This article is intended as
a \textit{practical guide} for researchers looking for in-depth help
with specifying FE models in SEM. It focuses on the \texttt{lavaan}
(Rosseel \protect\hyperlink{ref-R-lavaan}{2012}) package for \texttt{R}
(R Core Team \protect\hyperlink{ref-R-base}{2017}). While \texttt{Mplus}
(Muthén and Muthén, \protect\hyperlink{ref-Mplus}{n.d.}) is arguably the
most robust SEM software currently available (in terms of features like
alignment, latent variable interactions, for example), the
\texttt{lavaan} package has many benefits. First, like \texttt{R} it is
open source and completely free. For researchers dipping their toes into
SEM, there is no financial barrier to try, and no risk if they decide it
is not for them. Second, the implementation of \texttt{lavaan} in the
larger \texttt{R} environment is an enormous advantage. Instead of
poring over reams of plain text, copying out coefficients by hand, every
part of the \texttt{lavaan} output is available as an object. This means
that all aspects of the model, from fit indices, to coefficients and
standard errors, to the model matrices, can be accessed and easily
integrated into tables and plots. Furthermore, \texttt{R} can be used
for a great deal of applications. It can be used to manage and
manipulate as well as simulate data, perform symbolic algebra, run more
traditional analyses (e.g., multiple regression, logistic regression,
principal component analysis), etc. Once one is comfortable using
\texttt{R}, there is no longer any need to switch between different
software for data preparation and analysis.

The following article outlines the basic idea of FE regression, the
particularities of FE in SEM, and shows its implementation in
\texttt{lavaan}. Using
\href{https://github.com/henrik-andersen/FE-SEM/blob/master/simulation-code.R}{simulated
data}, it demonstrates and annotates the code for the most basic FE
model and provides an overview of the summary output. One of the main
strengths of panel SEM compared to the more traditional methods of panel
analysis is its flexibility. Therefore, a number of potential extensions
to the basic model, including relaxing various assumptions, dealing with
measurement error in both the independent and dependent variables, as
well as the inclusion of time-invariant predictors in the form of a
hybrid fixed-/ random effects model, are shown in detail in the form of
\href{https://github.com/henrik-andersen/FE-SEM/blob/master/extensions.pdf}{online
supplementary materials}.

\hypertarget{panel}{%
\section{Panel models}\label{panel}}

To begin, let us start by reviewing a general panel model (Bollen and
Brand \protect\hyperlink{ref-Bollen2010}{2010}), also referred to as the
`unobserved effects model' (Wooldridge
\protect\hyperlink{ref-Wooldridge2012}{2012}; Croissant and Millo
\protect\hyperlink{ref-R-plm_a}{2008}) (we will return to this model in
the
\href{https://github.com/henrik-andersen/FE-SEM/blob/master/extensions.pdf}{online
supplementary materials} of the article when we discuss loosening
assumptions) \begin{align}
y_{it} & = \bm{x}_{it}\bm{\beta} + \bm{z}_{i}\bm{\gamma} + \alpha_{i} + \varepsilon_{it} \label{eq:gpm}
\end{align} where \(y_{it}\) is the dependent variable for unit
\(i, \ i = 1, ..., N\) at time \(t, \ t = 1, ..., T\), \(\bm{x}_{it}\)
is a \(1 \times K\) vector of time-varying covariates (which could
include a constant) linked to the dependent variable by the
\(K \times 1\) vector of coefficients \(\bm{\beta}\). \(\bm{z}_{i}\) is
a \(1 \times M\) vector of time-invariant covariates linked to the
dependent variable by the \(M \times 1\) vector of coefficients in
\(\bm{\gamma}\), \(\alpha_{i}\) represents the combined effect of all
unobserved time-constant variables affecting the dependent variable and
\(\varepsilon_{it}\) is the idiosyncratic error.

We can make stating some of the model assumptions easier by rewriting it
in matrix notation \begin{align}
\bm{y}_{i} & = \bm{X}_{i}\bm{\beta} + \bm{Z}_{i}\bm{\gamma} + \bm{\iota}_{T}\alpha_{i} + \bm{\varepsilon}_{i}
\end{align} where \(\bm{y}_{i}\) and \(\bm{\varepsilon}_{i}\) are
\(T \times 1\) vectors, \(\bm{X}_{i}\) and \(\bm{Z}_{i}\) are
\(T \times K\) and \(T \times M\) matrices, respectively,
\(\bm{\iota}_{T}\) is a \(T \times 1\) vector of ones and \(\alpha_{i}\)
is a scalar. \(\bm{\beta}\) and \(\bm{\gamma}\) are unchanged from
Equation \eqref{eq:gpm}.

Consistency of the following models requires the assumption of strict
exogeneity, although what constitutes strict exogeneity differs between
the random and fixed effects setups. Each assumption will be discussed
shortly. Apart from that, we typically make the following assumptions
(see, e.g., Wooldridge \protect\hyperlink{ref-Wooldridge2002}{2002};
Schmidheiny \protect\hyperlink{ref-Schmidheiny2019}{2019}) about this
model:

\begin{itemize}
\tightlist
\item
  Linearity: the model is linear in its parameters.
\item
  Independence: the observations are independent across individuals
  (assured by random sampling in the cross-section), but not necessarily
  across time.
\item
  The usual rank condition: we have more observations than independent
  variables and there is no perfect collinearity between any of the
  independent variables.
\end{itemize}

\hypertarget{re}{%
\subsection{Random effects}\label{re}}

For the random effects (RE) model, we define a composite error term:
\(\nu_{it} = \alpha_{i} + \varepsilon_{it}\) and rewrite the model in
Equation \eqref{eq:gpm} as \begin{align}
y_{it} & = \bm{x}_{it}\bm{\beta} + \bm{z}_{i}\bm{\gamma} + \nu_{it}, \ \text{or} \\
\bm{y}_{i} & = \bm{X}_{i}\bm{\beta} + \bm{Z}_{i}\bm{\gamma} + \bm{\nu}_{i},
\end{align} where
\(\bm{\nu}_{i} = \alpha_{i} \bm{\iota}_{T} + \bm{\varepsilon}_{i}\) and
\(\bm{\iota}_{T}\) is a \(T \times 1\) vector of ones. The strict
exogeneity assumption in the RE model implies \begin{align}
\E[\varepsilon_{it} | \bm{X}_{i}, \bm{z}_{i}, \alpha_{i}] & = 0, \\
\E[\alpha_{i} |\bm{X}_{i}, \bm{z}_{i}] = \E[\alpha_{i}] & = 0,
\end{align} where \(\bm{X}_{i} = \bm{x}_{i1}, ..., \bm{x}_{iT}\). For
both parts, the assumption that the unconditional expectations are 0 is
unproblematic as long as a constant is included in the regression. The
first part says the idiosyncratic errors at each timepoint are assumed
to be independent of the explanatory variables at \textit{all}
timepoints which is stronger than just assuming that they are
\textit{contemporaneously} independent. This implies that they are also
uncorrelated, i.e.,
\(\E[\bm{x}_{is}^{\intercal}\varepsilon_{it}] = \bm{0}\) and
\(\E[\bm{z}_{i}^{\intercal}\varepsilon_{it}] = \bm{0}, \ \forall \ s, t = 1, ..., T\)
(Wooldridge \protect\hyperlink{ref-Wooldridge2002}{2002}; Brüderl and
Ludwig \protect\hyperlink{ref-Bruederl2015}{2015}). We assume the
idiosyncratic errors are further independent of the individual effects,
which implies \(\E[\alpha_{i}\varepsilon_{it}] = 0\).

The second part is the potentially controversial assumption: it states
that the individual effects are uncorrelated with the independent
variables. We can use an example to show why this is often
controversial: If we are interested in the question of whether married
men earn more than unmarried men, then the second part of the strict
exogeneity assumption means that a man's marriage status would have to
be uncorrelated with all the time-invariant characteristics that could
potentially make that man an attractive marriage candidate in the first
place; e.g., looks, personality, family's status, profession, etc.
(Brüderl and Ludwig
\protect\hyperlink{ref-Bruederl2015}{2015}).\footnote{Assuming, for the
  sake of argument, that these characteristics are constant over time.}

From what we have discussed so far, the \(T \times T\) covariance matrix
of the errors
\(\bm{\Omega}_{i} = \E[\bm{\nu}_{i}\bm{\nu}_{i}^{\intercal}]\) can be
constructed. However, the standard random effects model adds the
additional assumptions \begin{alignat}{3}
\E[\varepsilon_{it}^{2} | \bm{X}_{i}, \bm{z}_{i}, \alpha_{i}] & = \E[\varepsilon_{it}^{2}] && = \sigma_{\varepsilon}^{2}, && \ t = 1, ..., T, \\
\E[\varepsilon_{it}\varepsilon_{is}| \bm{X}_{i}, \bm{z}_{i}, \alpha_{i}] & = \E[\varepsilon_{it}\varepsilon_{is}] && = 0, && \ \forall \ t \ne s 
\end{alignat} i.e., the idiosyncratic errors are conditionally
homoscedastic and serially uncorrelated and \begin{alignat}{2}
\E[\alpha_{i}^{2}| \bm{X}_{i},\bm{z}_{i}] & = \E[\alpha_{i}^{2}] && = \sigma^{2}_{\alpha}
\end{alignat} i.e., the individual effects are conditionally
homoscedastic (they are necessarily serially correlated as long as
\(\sigma_{\alpha}^{2} > 0\)) we arrive at the typical random effects
structure of the \(NT \times NT\) matrix \(\bm{\Omega}\):

\begin{align}
\bm{\Omega} & = 
\begin{pmatrix}
\bm{\Omega}_{1} & \hdots & 0 & \hdots & 0 \\
\vdots & \ddots & \vdots & \ddots & \vdots \\
0 & \hdots & \bm{\Omega}_{i} & \hdots & 0 \\
\vdots & \ddots & \vdots & \ddots & \vdots \\
0 & \hdots & 0 & \hdots & \bm{\Omega}_{N}
\end{pmatrix}
\end{align} with \(T \times T\) typical elements \begin{align}
\bm{\Omega}_{i} & = 
\begin{pmatrix}
\sigma^{2}_{\nu} & \sigma^{2}_{\alpha} & \hdots & \sigma^{2}_{\alpha} \\
\sigma^{2}_{\alpha} & \sigma^{2}_{\nu} & \hdots & \sigma^{2}_{\alpha} \\
\vdots & \vdots & \ddots & \vdots \\
\sigma^{2}_{\alpha} & \sigma^{2}_{\alpha} & \hdots & \sigma^{2}_{\nu}
\end{pmatrix} \label{eq:omeganui}
\end{align} where
\(\sigma^{2}_{\nu} = \sigma^{2}_{\alpha} + \sigma^{2}_{\varepsilon}\)
(Wooldridge \protect\hyperlink{ref-Wooldridge2002}{2002}; Schmidheiny
\protect\hyperlink{ref-Schmidheiny2019}{2019}). This means that in the
conditional covariance matrix of the errors, given the time-varying and
-invariant covariates, units over time will be correlated due to the
individual effects. We should keep the covariance structure of the
errors in mind as it will help make sense of the use of latent variables
to decompose the dependent variable into between- and within-variance
components, discussed below in Section \ref{fe-sem}.

Estimation of the RE model can be done using feasible generalized least
squares (GLS) in which the two unknowns in \(\bm{\Omega}\),
\(\sigma_{\alpha}^{2}\) and \(\sigma_{\nu}^{2}\), are first estimated
using pooled ordinary least squares (pooled OLS or POLS),
where\footnote{Normally a degrees-of-freedom correction is applied by
  subtracting off the number of independent variables that is negligible
  in large N samples. It is ignored here for the sake of simplicity.}
\begin{align}
\hat{\sigma}_{\nu}^{2} & = \frac{1}{NT}\sum_{i = 1}^{N}\sum_{t = 1}^{T}\hat{\nu}_{it}^{2}, \\
\hat{\sigma}_{\varepsilon}^{2} & = \frac{1}{NT - N}\sum_{i = 1}^{N}\sum_{t = 1}^{T}(\hat{\nu}_{it} - \bar{\hat{\nu}}_{i})^{2}
\end{align} and
\(\hat{\nu}_{it} = y_{it} - \bm{x}_{it}\hat{\bm{\beta}}_{\tiny{POLS}} - \bm{z}_{i}\hat{\bm{\gamma}}_{\tiny{POLS}}\),
\(\bar{\hat{\nu}}_{i} = T^{-1}\sum_{t = 1}^{T}\hat{\nu}_{it}\) and
\(\hat{\sigma}^{2}_{\alpha} = \hat{\sigma}^{2}_{\nu} - \hat{\sigma}^{2}_{\varepsilon}\).
Then, the coefficients are estimated using those estimates in the
variance matrix \(\hat{\bm{\Omega}}\): \begin{align}
\begin{pmatrix}
\hat{\bm{\beta}}_{RE} \\
\hat{\bm{\gamma}}_{RE}
\end{pmatrix} & = 
(\bm{W}^{\intercal}\hat{\bm{\Omega}}^{-1}\bm{W})^{-1}\bm{W}^{\intercal}\hat{\bm{\Omega}}^{-1}\bm{y}
\end{align} where
\(\bm{W} = \begin{pmatrix}\bm{X} & \bm{Z}\end{pmatrix}\) and \(\bm{X}\)
is \(NT \times K\) and \(\bm{Z}\) is \(NT \times M\) and \(\bm{y}\) is
\(NT \times 1\).

In practice, however, computational problems can arise with large
cross-sectional samples, where it can become difficult to invert the
\(\hat{\bm{\Omega}}\) matrix. One solution is to use `partial-demeaning'
to transform the data before performing simple OLS: \begin{align}
(y_{it} - \theta\bar{y}_{i}) & = (\bm{x}_{it} - \theta\bar{\bm{x}}_{i})\bm{\beta} + (\bm{z}_{i} - \theta \bar{\bm{z}}_{i})\bm{\gamma} + (\varepsilon_{it} - \theta \bar{\varepsilon}_{i}) \label{eq:partialdemean}
\end{align} where
\(\theta = 1 - [\sigma_{\alpha}^{2}/(\sigma_{\alpha}^{2} + T \sigma_{\varepsilon}^{2})]^{1/2}\),
and \(\bar{y}_{i} = T^{-1}\sum_{t = 1}^{T}y_{it}\),
\(\bar{\bm{x}}_{i} = T^{-1}\sum_{t = 1}^{T}\bm{x}_{it}\),
\(\bar{\bm{z}}_{i} = T^{-1}\sum_{t = 1}^{T}\bm{z}_{i}\) and
\(\bar{\varepsilon}_{i} = T^{-1}\sum_{t = 1}^{T}\varepsilon_{it}\)
(Croissant and Millo \protect\hyperlink{ref-R-plm_a}{2008}).

The RE model can also be estimated in the maximum likelihood framework,
where in the associated literature panel models are generally referred
to as either mixed models, hierarchical models or longitudinal models.
The typical RE model discussed here is the equivalent to a mixed model
with random intercepts and fixed slopes (Croissant and Millo
\protect\hyperlink{ref-R-plm_a}{2008}). Under the assumption of
normality, along with homoscedasticity and serially uncorrelated errors,
the maximum likelihood estimator is the same as the OLS estimator. For
more on the topic of mixed models, see for example Bates et al.
(\protect\hyperlink{ref-R-lme4}{2015}), Bates
(\protect\hyperlink{ref-Bates2010}{2010}).

\hypertarget{fe}{%
\subsection{Fixed effects}\label{fe}}

For fixed effects, we assume that the individual effects are
\textit{not independent} of the model covariates, i.e.,
\(\E[\alpha_{i}|\bm{X}_{i},\bm{z}_{i}] \ne \E[\alpha_{i}] \ne 0\). Under
this assumption, grouping the individual effects in with the composite
error will cause the coefficients of interest, here specifically
\(\bm{\beta}\) to be inconsistent (Wooldridge
\protect\hyperlink{ref-Wooldridge2002}{2002},
\protect\hyperlink{ref-Wooldridge2012}{2012}). We write the model
therefore again in terms of the general panel model in Equation
\eqref{eq:gpm} with separate individual effects and idiosyncratic error
terms. In order to drop assumptions involving the individual effects, a
number of methods are available (e.g., differencing, least squares dummy
variable regression), but the most common approach is to \textit{demean}
the equation (Brüderl and Ludwig
\protect\hyperlink{ref-Bruederl2015}{2015}). Demeaning is the same as
the transformation applied in Equation \eqref{eq:partialdemean} in the
special case where \(\theta = 1\). I.e., demeaning involves subtracting
the per-unit, over-time average from each of the model terms, i.e.,
\begin{align}
(y_{it} - \bar{y}_{i}) & = (\bm{x}_{it} - \bar{\bm{x}}_{i})\bm{\beta} + (\bm{z}_{i} - \bar{\bm{z}}_{i})\bm{\gamma} + (\alpha_{i} - \bar{\alpha}_{i}) + (\varepsilon_{it} - \bar{\varepsilon}_{i}) \\
\ddot{y}_{it} & = \bm{\ddot{x}}_{it}\bm{\beta} + \ddot{\varepsilon}_{it}, \ \text{or} \\
\bm{\ddot{y}}_{i} & = \bm{\ddot{X}}_{i}\bm{\beta} + \bm{\ddot{\varepsilon}}_{i} \label{eq:fe}
\end{align} where the over-time averages are calculated the same as
above, and the variables with the dots above them represent the demeaned
versions. Because the average of something that does not change is that
thing itself, the individual effects, along with any time-invariant
predictors, get wiped out by the demeaning. This means that no
assumptions about the relatedness of the model covariates and the
unit-specific portion of the error are needed. Consistency of the
estimates is related solely to the strict exogeneity assumption imposed
on the idiosyncratic errors, i.e.,
\(\E[\ddot{\varepsilon}_{it}|\bm{\ddot{x}}_{it}] = \E[\ddot{\varepsilon}_{it}] = 0\)
which also implies
\(\E[\bm{\ddot{x}}_{is}^{\intercal}\ddot{\varepsilon}_{it}] = \bm{0}, \ \forall \ s, t = 1, ..., T\)
(Brüderl and Ludwig \protect\hyperlink{ref-Bruederl2015}{2015};
Wooldridge \protect\hyperlink{ref-Wooldridge2002}{2002}).

Having demeaned the data, the typical FE estimator is POLS on the
transformed data \begin{align}
\bm{\beta}_{FE} & = (\ddot{\bm{X}}^{\intercal}\ddot{\bm{X}})^{-1}\ddot{\bm{X}}^{\intercal}\ddot{\bm{y}}
\end{align} (Brüderl and Ludwig
\protect\hyperlink{ref-Bruederl2015}{2015}). The downside to this
approach is that no time-invariant predictors can be included in the
model. However, there are alternative approaches in the random effects
and mixed model frameworks that allow them to be included. These models
are sometimes referred to as `within-between' or `hybrid' models, often
based on the Chamberlain (\protect\hyperlink{ref-Chamberlain1980}{1980})
and Mundlak (\protect\hyperlink{ref-Mundlak1978}{1978}) approaches, see
for example Bell, Fairbrother, and Jones
(\protect\hyperlink{ref-Bell2018}{2018}); Allison
(\protect\hyperlink{ref-Allison2011}{2011}); Schunck
(\protect\hyperlink{ref-Schunck2013}{2013}); Enders and Tofighi
(\protect\hyperlink{ref-Enders2007}{2007}). In the
\href{https://github.com/henrik-andersen/FE-SEM/blob/master/extensions.pdf}{online
appendix}, it will be discussed how to also get around this restriction
using SEM.

\hypertarget{fe-sem}{%
\section{Fixed effects in structural equation modeling}\label{fe-sem}}

Moving from the conventional methods outlined above to SEM, we must
state the FE model in a different way. We turn to latent variables to
account for time-invariant unobserved heterogeneity. In fact, besides
accounting for measurement error and the representation of abstract
hypothetical concepts, unobserved heterogeneity has historically been
one of the main uses of latent variables in SEM (Skrondal and
Rabe-Hesketh \protect\hyperlink{ref-Skrondal2004}{2004}).

\hypertarget{modeling-time-invariant-unobserved-heterogeneity-as-a-latent-variable}{%
\subsection{Modeling time-invariant unobserved heterogeneity as a latent
variable}\label{modeling-time-invariant-unobserved-heterogeneity-as-a-latent-variable}}

We first need to convert the data from stacked, long-format vectors of
length \(NT\) into \(T\) individual vectors of length \(N\). To see why
this is necessary, consider what effect this has on the vector of
responses \(y_{it}\). Let us, for a minute ignore any covariates and
focus just on the dependent variable (a so-called `intercept-only' or
`null' model) so that we have
\(y_{it} = \alpha_{i} + \varepsilon_{it}\). When we convert the data to
wide-format, we get \(T\) individual equations, \begin{align}
\bm{y}_{t} & = \bm{\alpha} + \bm{\varepsilon}_{t} \\
\begin{bmatrix}
y_{1t} \\
y_{2t} \\
\vdots \\
y_{Nt}\end{bmatrix} & = 
\begin{bmatrix}
\alpha_{1} \\
\alpha_{2} \\
\vdots \\
\alpha_{N}
\end{bmatrix} + 
\begin{bmatrix}
\varepsilon_{1t} \\
\varepsilon_{2t} \\
\vdots \\
\varepsilon_{Nt}
\end{bmatrix} \label{eq:wide}
\end{align} for each \(t = 1, 2, ..., T\). Because the idiosyncratic
errors are assumed to be uncorrelated across units and across time, the
covariance between any two of the new wide vectors
\(\Cov(y_{ti},y_{si}) = \Var(\alpha_{i}), \ t \ne s\). Otherwise, when
\(t = s\), the covariance
\(\Cov(y_{ti},y_{ti}) = \Var(\alpha_{i}) + \Var(\varepsilon_{ti})\).
This is the structure we saw above in a typical element of
\(\bm{\Omega}\).

And in fact this is exactly how a latent variable is used to account for
time-invariant unobserved heterogeneity. The dependent variable at each
timepoint is regressed onto the latent variable, see Figure
\ref{fig:fesem}. Here, the regression weights or `factor loadings' are
fixed to one to represent our assumption that the effect of the
time-invariant unobserved heterogeneity is constant over time. It also
means that the estimated variance of the latent variable is equal to the
\textit{average covariance between the wide-format columns of the dependent variable over time}.
If \(y_{it} = \alpha_{i} + \varepsilon_{it}\) is the true data
generating process, then the relationship between two units over time is
just \(\alpha\), regardless of the time distance. Referring back to the
random effects structure of \(\bm{\Omega}_{i}\) in Equality
\eqref{eq:omeganui} for a generic unit \(i\), we see the covariance on
all of the off-diagonals is \(\sigma^{2}_{\alpha}\). And, as we know,
the average of something that does not change is that thing itself.
I.e., if \(T(T-1)/2 = h\) is the number of elements on either the upper-
or lower triangle of \(\bm{\Omega}_{i}\), then we have
\(h^{-1}\sum_{i=1}^{h}\sigma^{2}_{\alpha} = \frac{h \sigma^{2}_{\alpha}}{h} = \sigma^{2}_{\alpha}\).

To show this, consider the following matrix equation of the variances
and the nonredundant covariances in a three-wave intercept-only model
that follows directly from Equation \eqref{eq:wide} (assuming
\(\Cov(\varepsilon_{ti},\varepsilon_{si}) = 0, \ t \ne s\)), and which
we can solve easily with least squares: \begin{align}
\bm{A}\bm{x} & = \bm{b} \\
\begin{pmatrix}
1 & 1 & 0 & 0 \\
1 & 0 & 0 & 0 \\
1 & 0 & 0 & 0 \\
1 & 0 & 1 & 0 \\
1 & 0 & 0 & 0 \\
1 & 0 & 0 & 1
\end{pmatrix}
\begin{pmatrix}
\psi \\
\phi_{1} \\
\phi_{2} \\
\phi_{3}
\end{pmatrix} & = 
\begin{pmatrix}
\Var(y_{1}) \\
\Cov(y_{2},y_{1}) \\
\Cov(y_{3},y_{1}) \\
\Var(y_{2}) \\
\Cov(y_{3},y_{2}) \\
\Var(y_{3})
\end{pmatrix}
\end{align} where \(\psi = \Var(\alpha)\),
\(\phi_{t} = \Var(\varepsilon_{t})\). We can solve this equation to show
\begin{align}
\bm{A}\bm{x} & = \bm{b} \\
(\bm{A}^{\intercal}\bm{A})^{-1}\bm{A}^{\intercal}\bm{A}\bm{x} & = (\bm{A}^{\intercal}\bm{A})^{-1}\bm{A}^{\intercal}\bm{b} \\
\hat{\bm{x}} & = (\bm{A}^{\intercal}\bm{A})^{-1}\bm{A}^{\intercal}\bm{b} \\
\begin{pmatrix}
\hat{\psi} \\
\hat{\phi}_{1} \\
\hat{\phi}_{2} \\
\hat{\phi}_{3}
\end{pmatrix} & = 
\begin{pmatrix}
\frac{1}{3}\Cov(y_{2},y_{1}) + \frac{1}{3}\Cov(y_{3},y_{1}) + \frac{1}{3}\Cov(y_{3},y_{2}) \\
\Var(y_{1}) - \hat{\psi} \\
\Var(y_{2}) - \hat{\psi} \\
\Var(y_{3}) - \hat{\psi}
\end{pmatrix}.
\end{align} This shows that if our assumption of constant covariances
between the dependent variable across time holds, the estimated variance
of \(\alpha\) is just what it should be: the average covariance between
units of \(y\) over time. Once we add in observed covariates, the
estimated covariance of \(\alpha\) then become the \textit{conditional}
covariance of \(y\) over time, given those covariates.

\hypertarget{model-notation}{%
\subsection{Model notation}\label{model-notation}}

Having explained how a latent variable is used to estimate the
individual effects, we need to state the FE model using SEM-compatible
matrix notation. There are a number of different model notations (see,
for example Bollen (\protect\hyperlink{ref-Bollen1989}{1989}) for an
overview), but the one that will serve us best is one that was proposed
by Graff (\protect\hyperlink{ref-Graff1979}{1979}): \begin{align}
\bm{y}^{+} & = \bm{\Lambda_{y}}^{+} \bm{\eta}^{+}, \\
\bm{\eta}^{+} & = \bm{B}\bm{\eta}^{+} + \bm{\zeta}^{+}, 
\end{align} where
\(\bm{\eta}^{+} = (\bm{y}, \bm{x}, \bm{\eta}, \bm{\xi})^{\intercal}\),
\(\bm{\zeta}^{+} = (\bm{\varepsilon}, \bm{\delta}, \bm{\zeta}, \bm{\xi})^{\intercal}\),
\(\bm{y}^{+} = (\bm{y}, \bm{x})^{\intercal}\). \(\bm{y}\) is a vector of
observed dependent variables and \(\bm{x}\) is a vector of observed
independent variables. \(\bm{\eta}\) is a vector of the latent dependent
variables and \(\bm{\xi}\) is a vector of latent independent variables.
\(\bm{\varepsilon}\) and \(\bm{\delta}\) are vectors of the errors of
the observed dependent and independent variables, respectively, and
\(\bm{\zeta}\) is a vector of the errors, or disturbances, of the latent
variables. Notice the \(^{+}\) symbol is just meant to differentiate the
vectors with them from those without them. That means, \(\bm{\eta}^{+}\)
is a vector that holds the observed and latent variables, both dependent
(in SEM they are referred to as `endogenous') and independent (i.e.,
`exogeneous')\footnote{For our purposes, the terms endogenous and
  dependent, on the one hand, and exogenous and independent, on the
  other, can be used interchangeably.}, \(\bm{\zeta}^{+}\) holds the
errors for the observed variables and the disturbances of the latent
variables. \(\bm{y}^{+}\) holds just the observed variables, both
dependent and independent, and \(\bm{\Lambda_{y}}^{+}\) is a matrix of
ones and zeros that selects the observed variables from
\(\bm{\eta}^{+}\). Lastly, \(\bm{B}\) is a matrix that holds the
regression coefficients.

If we say that \(p\) and \(q\) stand for the number of observed
dependent and independent variables, respectively, and \(m\) and \(n\)
stand for the number of latent dependent and independent variables,
respectively, then \(\bm{\eta}^{+}\) and \(\bm{\zeta}^{+}\) are
\(p + q + m + n\), \(\bm{y}^{+}\) is \(p + q\), \(\bm{\Lambda_{y}}^{+}\)
is \((p + q) \times (p + q + m + n)\) and \(\bm{B}\) is
\((p + q + m + n) \times (p + q + m + n)\) (Bollen
\protect\hyperlink{ref-Bollen1989}{1989}).

This notation may be confusing at first, but it has advantages. First,
it allows us the flexibility we need for the models. For example, it
allows observed \(\bm{x}\) to directly influence observed \(\bm{y}\)
(more common notation assumes that substantive effects occur only
between latent variables, observed ones are only used as indicators, see
for example Bollen (\protect\hyperlink{ref-Bollen1989}{1989}), Kline
(\protect\hyperlink{ref-Kline2016}{2016})). It also allows \(\bm{\xi}\),
i.e., any latent exogenous variables, to influence \(\bm{y}\) directly.
These two scenarios cover the traditional FE model with observed
variables, and one in which latent variables are used to account for
measurement error in the independent variables. It is also consistent
with the notation used for these models in \texttt{lavaan}. In fact,
\texttt{lavaan} switches automatically between matrix notations
depending on the specified model. That means the matrix representation
of the model one sees if they type in
\texttt{lavInspect(model,\ what\ =\ "est")} after specifying their model
in \texttt{lavaan} will match up with the notation used here.\footnote{In
  \texttt{Mplus}, the model matrices can be requested by including
  \texttt{OUTPUT:\ TECH1} in the input file.} It does have a potential
disadvantage however. Besides being less intuitive than the typical
\(\bm{\eta} = \bm{B}\bm{\eta} + \bm{\Gamma}\bm{\xi} + \bm{\zeta}\)
notation, it means that by including the observed covariates in the
stacked long vector \(\bm{y}^{+}\), they are treated as another response
variable with variances and covariances to be estimated by the model
(instead of just using the sample statistics). This means that the
assumption of multivariate normality (otherwise just imposed on the
dependent variables) also applies to the independent ones (Skrondal and
Rabe-Hesketh \protect\hyperlink{ref-Skrondal2004}{2004}, 75). This can
be problematic for noncontinuous independent variables like sex,
nationality dummies, marriage status (married/unmarried), etc. See
Skrondal and Rabe-Hesketh (\protect\hyperlink{ref-Skrondal2004}{2004})
for more on this topic.

\begin{figure}
\begin{center}
\resizebox{0.75\textwidth}{!}{%
\begin{tikzpicture}
% Node styles ---
\tikzstyle{man} = [rectangle, thick, minimum size = 1cm, draw = black!80, fill = white!100, font = \sffamily] % Manifest variables
\tikzstyle{lat} = [circle, thick, minimum size = 1cm, draw = black!80, fill = white!100, font = \sffamily] % Latent variables
\tikzstyle{err} = [circle, draw = black!80, fill = white!100, font = \sffamily] % Errors 
% Edge styles
\tikzstyle{con} = [-latex, font = \sffamily] % Effects
\tikzstyle{cons} = [-latex, font = \sffamily\small] % Effects with smaller font 
\tikzstyle{cor} = [latex-latex, font = \sffamily] % Correlations
% Begin Figure ---
% Nodes 
\node at (+0.0,+0.0) [man] (x1) {$x_{1}$};
\node at (+2.0,+0.0) [man] (x2) {$x_{2}$};
\node at (+6.0,+0.0) [man] (x3) {$x_{3}$};
\node at (+4.0,+0.0) [lat] (eta) {$\alpha$};
\node at (+0.0,-2.0) [man] (y1) {$y_{1}$};
\node at (+2.0,-2.0) [man] (y2) {$y_{2}$};
\node at (+6.0,-2.0) [man] (y3) {$y_{3}$};
\node at (+0.0,-3.5) [err] (e1) {$\varepsilon_{1}$};
\node at (+2.0,-3.5) [err] (e2) {$\varepsilon_{2}$};
\node at (+6.0,-3.5) [err] (e3) {$\varepsilon_{3}$};
% Paths
\path (x1) edge [con, draw = black!100, left] node {$\beta$} (y1);
\path (x2) edge [con, draw = black!100, left, near start] node {$\beta$} (y2);
\path (x3) edge [con, draw = black!100, right] node {$\beta$} (y3);
\path (e1) edge [con, draw = black!100, left] node {1} (y1);
\path (e2) edge [con, draw = black!100, left] node {1} (y2);
\path (e3) edge [con, draw = black!100, right] node {1} (y3);
\path (eta) edge [con, draw = black!100, above, near start] node {1} (y1);
\path (eta) edge [con, draw = black!100, below] node {1} (y2);
\path (eta) edge [con, draw = black!100, above] node {1} (y3);
% correlations
\path (x1) edge [cor, bend left] node {} (x2);
\path (x1) edge [cor, bend left] node {} (eta);
\path (x1) edge [cor, bend left] node {} (x3);
\path (x2) edge [cor, bend left] node {} (eta);
\path (x2) edge [cor, bend left] node {} (x3);
\path (x3) edge [cor, bend right] node {} (eta);
\end{tikzpicture}
}
\caption{Typical three-wave FE-SEM model with contemporary effects\label{fig:fesem}}
\end{center}
\end{figure}

Let us, however, make things more concrete and take a look at a simple,
three-wave version of the typical FE-SEM using this notation (shown
graphically in Figure \ref{fig:fesem}). For that, we have the following
matrix notation (with labels on the outside of the matrices):
\begin{align}
\begin{split}
\bm{y}^{+} & = \bm{\Lambda_{y}}^{+}\bm{\eta}^{+} \\
\begin{pmatrix}
y_{1} \\
y_{2} \\
y_{3} \\
x_{1} \\
x_{2} \\
x_{3}
\end{pmatrix} & = 
\begin{blockarray}{cccccccc}
 & y_{1} & y_{2} & y_{3} & x_{1} & x_{2} & x_{3} & \alpha \\
 \begin{block}{c(ccccccc)}
 y_{1} & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
 y_{2} & 0 & 1 & 0 & 0 & 0 & 0 & 0 \\
 y_{3} & 0 & 0 & 1 & 0 & 0 & 0 & 0 \\ 
 x_{1} & 0 & 0 & 0 & 1 & 0 & 0 & 0 \\
 x_{2} & 0 & 0 & 0 & 0 & 1 & 0 & 0 \\
 x_{3} & 0 & 0 & 0 & 0 & 0 & 1 & 0 \\
 \end{block}
\end{blockarray}
\begin{pmatrix}
y_{1} \\
y_{2} \\
y_{3} \\
x_{1} \\
x_{2} \\
x_{3} \\
\alpha
\end{pmatrix},
\end{split} \label{eq:y}
\end{align} \begin{align}
\begin{split}
\bm{\eta}^{+} & = \bm{B}\bm{\eta}^{+} + \bm{\zeta}^{+} \\
\begin{pmatrix}
y_{1} \\
y_{2} \\
y_{3} \\
x_{1} \\
x_{2} \\
x_{3} \\
\alpha
\end{pmatrix} & = 
\begin{blockarray}{cccccccc}
 & y_{1} & y_{2} & y_{3} & x_{1} & x_{2} & x_{3} & \alpha \\
 \begin{block}{c(ccccccc)}
 y_{1}  & 0 & 0 & 0 & \beta & 0 & 0 & 1 \\
 y_{2}  & 0 & 0 & 0 & 0 & \beta & 0 & 1\\
 y_{3}  & 0 & 0 & 0 & 0 & 0 & \beta & 1 \\ 
 x_{1}  & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
 x_{2}  & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
 x_{3}  & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
 \alpha & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
 \end{block}
\end{blockarray}
\begin{pmatrix}
y_{1} \\
y_{2} \\
y_{3} \\
x_{1} \\
x_{2} \\
x_{3} \\
\alpha
\end{pmatrix} + 
\begin{pmatrix}
\varepsilon_{1} \\
\varepsilon_{2} \\
\varepsilon_{3} \\
x_{1} = \delta_{1} \\
x_{2} = \delta_{2} \\
x_{3} = \delta_{3} \\
\alpha = \xi \\
\end{pmatrix}. 
\end{split} \label{eq:eta}
\end{align} Notice in \(\bm{\zeta}^{+}\), for the independent variables
we could either write, for example \(x_{t}\) or \(\delta_{t}\). As
mentioned above, this is due to the model notation treating the
independent variables like dependent variables with
variances/covariances to be estimated. For the sake of simplicity, we
will ignore this subtlety and refer to the observed variable from now
on, keeping in mind that if the multivariate normality assumption holds,
the estimated statistics will likely be sufficiently close to the sample
ones for it to not make much of a difference.

Admittedly, Equations \eqref{eq:y} and \eqref{eq:eta} may not look like
much yet. We can remedy this by first putting the equation for
\(\bm{\eta}^{+}\) in reduced form, i.e., by getting rid of the dependent
variable on the r.h.s.: \begin{align}
\bm{\eta}^{+} & = \bm{B}\bm{\eta}^{+} + \bm{\zeta}^{+} \\
\bm{\eta}^{+} - \bm{B}\bm{\eta}^{+} & = \bm{\zeta}^{+} \\
(\bm{I} - \bm{B})\bm{\eta}^{+} & = \bm{\zeta}^{+} \\
\bm{\eta}^{+} & = (\bm{I} - \bm{B})^{-1}\bm{\zeta}^{+},
\end{align} where \(\bm{I}\) is the identity matrix. By substituting
this back into the equation for the observed variables we get
\(\bm{y}^{+} = \bm{\Lambda_{y}}^{+}[(\bm{I} - \bm{B})^{-1}\bm{\zeta}^{+}]\),
which works out to: \begin{align}
\begin{pmatrix} 
y_{1} \\
y_{2} \\
y_{3} \\
x_{1} \\
x_{2} \\
x_{3} \\
\end{pmatrix} & = 
\begin{pmatrix}
\alpha + \beta x_{1} + \varepsilon_{1} \\
\alpha + \beta x_{2} + \varepsilon_{2} \\
\alpha + \beta x_{3} + \varepsilon_{3} \\
x_{1} \\
x_{2} \\
x_{3}
\end{pmatrix},
\end{align} which is of course exactly what we should expect given
Equation \eqref{eq:fe}\footnote{We can use the \texttt{sympy} package in
  \texttt{python} to verify and show the steps for this and other
  examples, see the
  \href{https://github.com/henrik-andersen/FE-SEM/blob/master/sympy-doublecheck-matrixnotation.py}{supplementary
  materials} for the code.}.

\hypertarget{assumptions}{%
\subsection{Assumptions}\label{assumptions}}

What essentially differentiates an FE from a random effects (RE) model
is our assumption concerning the relationship between the unobserved
individual effects and the model covariates (Bollen and Brand
\protect\hyperlink{ref-Bollen2010}{2010}). The FE model assumes that
\(\E[\alpha x_{t}] \ne 0\). As such, if we fail to control for the
correlation of the covariate and the time-invariant part of the error,
then the coefficient of interest, here \(\beta\), will be biased. Our
assumption regarding whether the individual effects are correlated with
the model covariates occurs in
\(\E[\bm{\zeta}^{+}\bm{\zeta}^{+ \intercal}] = \bm{\Psi}\), the
covariance matrix of the errors \begin{align}
\bm{y}^{+}\bm{y}^{+ \intercal} & = \E[(\bm{\Lambda_{y}}^{+}(\bm{I} - \bm{B})^{-1}\bm{\zeta}^{+})(\bm{\Lambda_{y}}^{+}(\bm{I} - \bm{B})^{-1}\bm{\zeta}^{+})^{\intercal}] \\
 & = \E[(\bm{\Lambda_{y}}^{+}(\bm{I} - \bm{B})^{-1}\bm{\zeta}^{+})(\bm{\zeta}^{+ \intercal}(\bm{I} - \bm{B})^{-1 \intercal}\bm{\Lambda_{y}}^{+ \intercal})] \\
 & = \bm{\Lambda_{y}}^{+}(\bm{I} - \bm{B})^{-1} \E[\bm{\zeta}^{+}\bm{\zeta}^{+ \intercal}] (\bm{I} - \bm{B})^{-1 \intercal}\bm{\Lambda_{y}}^{+ \intercal} \\
 & = \bm{\Lambda_{y}}^{+}(\bm{I} - \bm{B})^{-1} \bm{\Psi} (\bm{I} - \bm{B})^{-1 \intercal}\bm{\Lambda_{y}}^{+ \intercal}.
\end{align} In the case of an FE model, \(\bm{\Psi}\) will reflect our
belief that the individual effects are correlated with the model
covariates, here again for demonstration the three-wave model:
\begin{align}
\bm{\Psi} & = \E
\begin{blockarray}{cccccccc}
 & \varepsilon_{1} & \varepsilon_{2} & \varepsilon_{3} & x_{1} & x_{2} & x_{3} & \alpha \\
 \begin{block}{c(ccccccc)}
 \varepsilon_{1} & \varepsilon_{1}^{2} &                     &                     &              &              &              & \\
 \varepsilon_{2} & 0                   & \varepsilon_{2}^{2} &                     &              &              &              & \\
 \varepsilon_{3} & 0                   & 0                   & \varepsilon_{3}^{2} &              &              &              & \\
 x_{1}           & 0                   & 0                   & 0                   & x_{1}^{2}    &              &              & \\
 x_{2}           & 0                   & 0                   & 0                   & x_{2}x_{1}   & x_{2}^{2}    &              & \\
 x_{3}           & 0                   & 0                   & 0                   & x_{3}x_{1}   & x_{3}x_{2}   & x_{3}^{2}    & \\
 \alpha          & 0                   & 0                   & 0                   & \alpha x_{1} & \alpha x_{2} & \alpha x_{3} & \alpha^{2} \\
 \end{block}
\end{blockarray}.
\end{align} Knowing this, we can work out the equation for the
coefficient of interest, \(\beta\). For the sake of simplicity, assume
here and throughout mean-centered variables: \begin{align}
\Cov(y_{t},x_{t}) & = \E[y_{t}x_{t}] \\
 & = \E[(\alpha + \beta x_{t} + \varepsilon_{t})x_{t}] \\
 & = \E[\alpha x_{t} + \beta x_{t}^{2} + \varepsilon_{t}x_{t}] \\
 & = \Cov(\alpha, x_{t}) + \beta \Var(x_{t}) \\
\hat{\beta}_{FE-SEM} & = \frac{\Cov(y_{t},x_{t}) - \Cov(\alpha, x_{t})}{\Var(x_{t})}. 
\end{align} This should make intuitive sense. From the observed
covariance between the dependent and the independent variable, we are
partialling out the part that is due to the covariance between the
independent variable and the individual effects per unit, and then
dividing by the variance of the independent variable, as usual. For the
RE model, we assume \(\E[\alpha x_{t}] = 0\) and the equation reduces to
\(\hat{\beta}_{RE-SEM} = \Cov(y_{t},x_{t})/\Var(x_{t})\). The rest of
the model-implied covariance matrix results from
\(\bm{y}^{+}\bm{y}^{+ \intercal}\).

\hypertarget{fesem}{%
\section{Fixed effects in lavaan}\label{fesem}}

The package \texttt{lavaan} needs to be installed once with
\texttt{install.packages("lavaan")}. To be able to use it, we need to
load it for every new \texttt{R} session:

\singlespacing

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{( lavaan)}
\end{Highlighting}
\end{Shaded}

\doublespacing

For users unfamiliar with \texttt{R}, SEM analyses can be carried out
with almost no knowledge of the language. Typically, someone unfamiliar
with \texttt{R} would prepare their data using some other statistical
software, and then save the intended dataset as a \texttt{.csv},
\texttt{.xlsx}, \texttt{.dta}, \texttt{.sav}, etc. file. The user must
then import the data, preferably as a dataframe, and the rest occurs
using the \texttt{lavaan} syntax.\footnote{There are many online
  tutorials for importing data in various formats, see, for example some
  from
  \href{https://www.datacamp.com/community/tutorials/r-data-import-tutorial}{datacamp}
  or
  \href{https://www.statmethods.net/input/importingdata.html}{Quick-R},
  or any of the many posts on
  \href{https://stackoverflow.com/search?q=r+import+data}{stackoverflow}.}

Specifying the most basic fixed effects model, like the one shown in
Bollen and Brand (\protect\hyperlink{ref-Bollen2010}{2010}) (the same
model as Equation \eqref{eq:fe} but with just one time-varying
predictor) involves four components. First, we define the latent
individual effects variable using the \texttt{=\textasciitilde{}}
`measured by' or `manifested by' (Rosseel
\protect\hyperlink{ref-R-lavaan}{2012}) operator at the same time
constraining the factor loadings at each timepoint to one. I will call
the latent variable \texttt{a} to stand for \(\alpha\). Constraining all
of the factor loadings to one reflects our implicit assumption that the
combined effect of the unit-specific unobserved factors is constant over
time. This is the default behaviour of traditional POLS-based approaches
to FE that use the stacked long-format data.

\singlespacing

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{a =}\ErrorTok{~}\StringTok{ }\DecValTok{1}\OperatorTok{*}\NormalTok{y1 }\OperatorTok{+}\StringTok{ }\DecValTok{1}\OperatorTok{*}\NormalTok{y2 }\OperatorTok{+}\StringTok{ }\DecValTok{1}\OperatorTok{*}\NormalTok{y3 }\OperatorTok{+}\StringTok{ }\DecValTok{1}\OperatorTok{*}\NormalTok{y4 }\OperatorTok{+}\StringTok{ }\DecValTok{1}\OperatorTok{*}\NormalTok{y5}
\end{Highlighting}
\end{Shaded}

\doublespacing

Second, we regress the dependent variable on the independent variable
using the \texttt{\textasciitilde{}} regression operator. With stacked,
long-format data, only one regression coefficient is estimated over all
observed timepoints. To have our FE-SEM model mimic this behaviour, we
need to constrain the the estimated coefficient to equal over time. We
do so by adding the same label to the regression coefficient at every
time point. We will use the label \texttt{b} (this label was chosen
arbitrarily, we could have used any letter or string of characters) and
have it act as an equality constraint for the regression coefficient of
interest \(\beta\):

\singlespacing

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{y1 }\OperatorTok{~}\StringTok{ }\NormalTok{b}\OperatorTok{*}\NormalTok{x1}
\NormalTok{y2 }\OperatorTok{~}\StringTok{ }\NormalTok{b}\OperatorTok{*}\NormalTok{x2 }
\NormalTok{y3 }\OperatorTok{~}\StringTok{ }\NormalTok{b}\OperatorTok{*}\NormalTok{x3}
\NormalTok{y4 }\OperatorTok{~}\StringTok{ }\NormalTok{b}\OperatorTok{*}\NormalTok{x4}
\NormalTok{y5 }\OperatorTok{~}\StringTok{ }\NormalTok{b}\OperatorTok{*}\NormalTok{x5}
\end{Highlighting}
\end{Shaded}

\doublespacing

The key to a FE model, as opposed to an RE model are our assumptions
about the relatedness of our covariate and the individual effects, i.e.,
\(\E[x_{t}\alpha]\). For an FE model, we want to partial out any
potential covariance between the independent variable and the individual
effects. This accounts for any linear relationship between \(x_{t}\) and
the unit-specific characteristics influencing the dependent variable.
Further, allowing unrestricted covariances between the independent
variable itself over time will not affect how the coefficient \(\beta\)
is estimated, but will have an effect on the standard errors. To mimic
the behaviour of a conventional FE model, we allow the independent
variable to be correlated with the individual effects and itself over
time. Covariances (including covariances between a variable and itself,
i.e., variances) are specified using the
\texttt{\textasciitilde{}\textasciitilde{}} operator:

\singlespacing

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{a }\OperatorTok{~}\ErrorTok{~}\StringTok{ }\NormalTok{x1 }\OperatorTok{+}\StringTok{ }\NormalTok{x2 }\OperatorTok{+}\StringTok{ }\NormalTok{x3 }\OperatorTok{+}\StringTok{ }\NormalTok{x4 }\OperatorTok{+}\StringTok{ }\NormalTok{x5}
\NormalTok{x1 }\OperatorTok{~}\ErrorTok{~}\StringTok{ }\NormalTok{x2 }\OperatorTok{+}\StringTok{ }\NormalTok{x3 }\OperatorTok{+}\StringTok{ }\NormalTok{x4 }\OperatorTok{+}\StringTok{ }\NormalTok{x5}
\NormalTok{x2 }\OperatorTok{~}\ErrorTok{~}\StringTok{ }\NormalTok{x3 }\OperatorTok{+}\StringTok{ }\NormalTok{x4 }\OperatorTok{+}\StringTok{ }\NormalTok{x5}
\NormalTok{x3 }\OperatorTok{~}\ErrorTok{~}\StringTok{ }\NormalTok{x4 }\OperatorTok{+}\StringTok{ }\NormalTok{x5}
\NormalTok{x4 }\OperatorTok{~}\ErrorTok{~}\StringTok{ }\NormalTok{x5}
\end{Highlighting}
\end{Shaded}

\doublespacing

The last component of our code involves the variances of the residuals.
This component is optional, but we can constrain the residual variances
to be equal over time to again mimic the behaviour of a conventional FE
model using POLS on stacked data. Here, again, we use labels to make
equality constraints. Because \(y_{t}\) is endogenous, the
\texttt{\textasciitilde{}\textasciitilde{}} operator specifies the
variances of \emph{residuals}, i.e., \(\varepsilon_{t}\).

\singlespacing

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{y1 }\OperatorTok{~}\ErrorTok{~}\StringTok{ }\NormalTok{e}\OperatorTok{*}\NormalTok{y1}
\NormalTok{y2 }\OperatorTok{~}\ErrorTok{~}\StringTok{ }\NormalTok{e}\OperatorTok{*}\NormalTok{y2}
\NormalTok{y3 }\OperatorTok{~}\ErrorTok{~}\StringTok{ }\NormalTok{e}\OperatorTok{*}\NormalTok{y3}
\NormalTok{y4 }\OperatorTok{~}\ErrorTok{~}\StringTok{ }\NormalTok{e}\OperatorTok{*}\NormalTok{y4}
\NormalTok{y5 }\OperatorTok{~}\ErrorTok{~}\StringTok{ }\NormalTok{e}\OperatorTok{*}\NormalTok{y5}
\end{Highlighting}
\end{Shaded}

\doublespacing

\hypertarget{ex1}{%
\section{A simulated example}\label{ex1}}

\singlespacing

\doublespacing

To demonstrate the application of FE models in SEM, a dataset can be
simulated that embodies the FE assumptions. Again, the code for data
simulation can be found in the
\href{https://github.com/henrik-andersen/FE-SEM/blob/master/simulation-code.R}{online
supplementary materials}.

To show that the latent individual effects variables represent the
\emph{combined} effect of all time-invariant characteristics, the
dependent variable will be influenced by two separate unit-specific
variables, which we can call \(\alpha_{1}\) and \(\alpha_{2}\). We will
construct the simulated data such that the independent variable is
correlated with both of the time-invariant variables. This means that
approaches that fail to account for this confounding influence, such as
POLS or RE, will be biased.

The wide-format equations for the data generating process can be
described as: \begin{align}
\bm{x_{t}} & = \bm{\alpha_{1}}\beta_{x_{t},\alpha_{1}} + \bm{\alpha_{2}}\beta_{x_{t},\alpha_{2}} + \bm{\delta_{t}}, \\
\bm{y_{t}} & = \bm{x_{t}}\beta_{y_{t},x_{t}} + \bm{\alpha_{1}}\beta_{y_{t},\alpha_{1}} + \bm{\alpha_{2}} \beta_{y_{t},\alpha_{2}} + \bm{\varepsilon_{t}} 
\end{align} where, for the sake of simplicity, \(\bm{\alpha_{1}}\),
\(\bm{\alpha_{2}}\), \(\bm{\delta_{t}}\) and \(\bm{\varepsilon_{t}}\)
are \(\sim N(0,1)\).

For the following example, a sample size of 1,000, observed over five
waves, was chosen. The unique variance of \(\bm{x}\), as well as both
the individual-effect variables is also \(\sim N(0,1)\). The coefficient
of interest, \(\beta_{y,x}\) is set to be equal to \(0.3\). A
correlation between \(\bm{x}\) and the individual effects is induced
through \(\beta_{x,\alpha_{1}} = 0.85\) and
\(\beta_{x,\alpha_{1}} = 0.50\). With the variances above set to one,
the covariances will be roughly \(\Cov(x_{t},\alpha_{1}) = 0.85\) and
\(\Cov(x_{t},\alpha_{2}) = 0.5\). The dependent variable is also
influenced by the individual effects variables with
\(\beta_{y_{t},\alpha_{1}} = 0.75\) and
\(\beta_{y_{t},\alpha_{2}} = 0.45\) These values were chosen
arbitrarily.

Now, we run the FE-SEM in \texttt{lavaan}.

\singlespacing

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{fe_sem <-}\StringTok{ '}
\StringTok{# Define individual effects variable }
\StringTok{a =~ 1*y1 + 1*y2 + 1*y3 + 1*y4 + 1*y5}
\StringTok{# Regressions, constrain coefficient to be equal over time}
\StringTok{y1 ~ b*x1}
\StringTok{y2 ~ b*x2 }
\StringTok{y3 ~ b*x3}
\StringTok{y4 ~ b*x4}
\StringTok{y5 ~ b*x5}
\StringTok{# Allow unrestricted correlation between eta and covariates}
\StringTok{a ~~ x1 + x2 + x3 + x4 + x5}
\StringTok{x1 ~~ x2 + x3 + x4 + x5}
\StringTok{x2 ~~ x3 + x4 + x5}
\StringTok{x3 ~~ x4 + x5}
\StringTok{x4 ~~ x5}
\StringTok{# Constrain residual variances to be equal over time}
\StringTok{y1 ~~ e*y1}
\StringTok{y2 ~~ e*y2}
\StringTok{y3 ~~ e*y3}
\StringTok{y4 ~~ e*y4}
\StringTok{y5 ~~ e*y5}
\StringTok{'}
\NormalTok{fe_sem.fit <-}\StringTok{ }\KeywordTok{sem}\NormalTok{(}\DataTypeTok{model =}\NormalTok{ fe_sem, }
                  \DataTypeTok{data =}\NormalTok{ dfw, }
                  \DataTypeTok{estimator =} \StringTok{"ML"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\doublespacing

\singlespacing

\doublespacing

We can get a summary of the model with \texttt{summary()}. The first
portion of the summary output gives an overview of some basic
information and fit statistics. The maximum likelihood estimator is the
default, so it did not have to be explicitly selected in the fitting
function call. Other estimators are available, including generalized and
unweighted least squares (\texttt{GLS} and \texttt{ULS}, respectively),
robust standard errors maximum likelihood (\texttt{MLM}) and several
others (see \href{https://lavaan.ugent.be/tutorial/est.html}{the lavaan
online tutorial for more}).

This part of the summary output also tells us that the analysis is based
on 1,000 observations (missings would be shown here as well if there
were any), and that the \(\chi^{2}\) statistic is 30.138 based on 32
degrees of freedom (55 observed covariances minus 1 error variance, 1
coefficient, 1 latent variable variance, 5 exogenous variable variances
and 15 covariances for \(55 - 23 = 32\) df). The p-value on the
\(\chi^{2}\) statistic is not significant with \(p =\) 0.561 which tells
us the differences between the model-implied and observed covariance
matrices are likely due to chance, and that the model fits the data well
(given how the data was generated, it would be surprising if this were
not the case). Other fit measures including typical comparative fit
indices can be requested by either adding \texttt{fit.measures\ =\ TRUE}
as a secondary argument to the \texttt{summary()} call, or by asking for
a complete list of all available fit statistics using
\texttt{lavInspect(model,\ "fit")} where \texttt{model} stands for the
name of the fitted model, in this case \texttt{fe\_sem.fit}.

\singlespacing

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{summary}\NormalTok{(fe_sem.fit)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## lavaan 0.6-6 ended normally after 37 iterations
## 
##   Estimator                                         ML
##   Optimization method                           NLMINB
##   Number of free parameters                         31
##   Number of equality constraints                     8
##                                                       
##   Number of observations                          1000
##                                                       
## Model Test User Model:
##                                                       
##   Test statistic                                30.138
##   Degrees of freedom                                32
##   P-value (Chi-square)                           0.561
## 
## Parameter Estimates:
## 
##   Standard errors                             Standard
##   Information                                 Expected
##   Information saturated (h1) model          Structured
...
\end{verbatim}

\doublespacing

Next the summary output shows the measurement models for the latent
variables, if any. In this case the latent variable \texttt{a} for
\(\alpha\) is measured by each of the five observed dependent variables
with factor loadings fixed to 1.0.

\singlespacing

\begin{verbatim}
...
## Latent Variables:
##                    Estimate  Std.Err  z-value  P(>|z|)
##   a =~                                                
##     y1                1.000                           
##     y2                1.000                           
##     y3                1.000                           
##     y4                1.000                           
##     y5                1.000                           
...
\end{verbatim}

\doublespacing

The regressions are shown next. Here, because we have constrained the
regression coefficients to be equal over time (the equality constraint
label \texttt{(b)} is listed to the left of the estimates), the estimate
of \(\beta =\) 0.294 (0.016) is repeated five times. The corresponding
z- and p-values show that the coefficient is, unsurprisingly,
significant.

\singlespacing

\begin{verbatim}
...
## Regressions:
##                    Estimate  Std.Err  z-value  P(>|z|)
##   y1 ~                                                
##     x1         (b)    0.294    0.016   18.809    0.000
##   y2 ~                                                
##     x2         (b)    0.294    0.016   18.809    0.000
##   y3 ~                                                
##     x3         (b)    0.294    0.016   18.809    0.000
##   y4 ~                                                
##     x4         (b)    0.294    0.016   18.809    0.000
##   y5 ~                                                
##     x5         (b)    0.294    0.016   18.809    0.000
...
\end{verbatim}

\doublespacing

Next, the covariance estimates are listed. First, the covariances
between the latent individual effects variable and the independent
variable over time are shown, and then the covariances between the
independent variable with itself over time.

One should always take care to double-check that there are no unintended
covariances listed here. Like \texttt{Mplus}, the \texttt{lavaan}
package estimates some covariances per default, without the user
explicitly having to add them to the model syntax. For example,
covariances between latent variables are estimated per default. If one
does not wish for them to covary, it must be explicitly stated, e.g.,
with \texttt{f1\ \textasciitilde{}\textasciitilde{}\ 0*f2}, assuming the
latent variables are called \texttt{f1} and \texttt{f2}, or by
overriding the default behaviour for the entire model by adding
\texttt{orthogonal\ =\ TRUE} (which sets the correlation between all
latent variables to zero) to the fitting call.\footnote{This is at least
  the current behaviour of both the \texttt{cfa} and \texttt{sem}
  wrappers. In fact, both wrappers seem to be identical in terms of the
  default settings, see Rosseel et al.
  (\protect\hyperlink{ref-Rosseel2020}{2020}).}

\singlespacing

\begin{verbatim}
...
## Covariances:
##                    Estimate  Std.Err  z-value  P(>|z|)
##   a ~~                                                
##     x1                0.844    0.055   15.355    0.000
##     x2                0.867    0.056   15.441    0.000
##     x3                0.845    0.055   15.400    0.000
##     x4                0.822    0.053   15.455    0.000
##     x5                0.820    0.053   15.572    0.000
##   x1 ~~                                               
##     x2                0.908    0.070   12.900    0.000
##     x3                0.935    0.069   13.466    0.000
##     x4                0.921    0.067   13.661    0.000
##     x5                0.914    0.067   13.716    0.000
##   x2 ~~                                               
##     x3                0.889    0.070   12.675    0.000
##     x4                0.922    0.069   13.423    0.000
##     x5                0.889    0.068   13.165    0.000
##   x3 ~~                                               
##     x4                0.865    0.067   12.976    0.000
##     x5                0.901    0.066   13.554    0.000
##   x4 ~~                                               
##     x5                0.850    0.064   13.285    0.000
...
\end{verbatim}

\doublespacing

Finally, the variance estimates are listed. Here, we see that in order
to mimic the behaviour of a traditional FE model, the error variances
over time were specified to be equal using the equality constraint
\texttt{(e)}. Notice the \texttt{.} beside \texttt{y1}, \texttt{y2},
etc.: this indicates that the listed variance refers to an endogenous
variable, and that it is thus an error variance. In this case, these
refer to the variances of \(\varepsilon_{t}\). After that, the variances
of the exogenous variables, both observed and unobserved are listed.

\singlespacing

\begin{verbatim}
...
## Variances:
##                    Estimate  Std.Err  z-value  P(>|z|)
##    .y1         (e)    1.022    0.023   44.721    0.000
##    .y2         (e)    1.022    0.023   44.721    0.000
##    .y3         (e)    1.022    0.023   44.721    0.000
##    .y4         (e)    1.022    0.023   44.721    0.000
##    .y5         (e)    1.022    0.023   44.721    0.000
##     x1                1.986    0.089   22.361    0.000
##     x2                2.079    0.093   22.361    0.000
##     x3                1.987    0.089   22.361    0.000
##     x4                1.860    0.083   22.361    0.000
##     x5                1.814    0.081   22.361    0.000
##     a                 0.799    0.052   15.310    0.000
\end{verbatim}

\doublespacing

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

Fixed effects regression in SEM has been outlined in well-known articles
by (Allison \protect\hyperlink{ref-Allison2011}{2011}; Bollen and Brand
\protect\hyperlink{ref-Bollen2010}{2010}; Teachman et al.
\protect\hyperlink{ref-Teachman2001}{2001}). This article provides a
focused look at the implementation of the basic model, as well as common
extensions using the \texttt{lavaan} package in \texttt{R}.

The benefits of FE-SEM as opposed to traditional OLS-based FE-models are
largely the same ones that apply to the SEM framework in general: for
one, SEM allows for a great deal of flexibility. For example, it is easy
to loosen model constraints as necessary. Measurement error in both the
dependent and independent variables can be dealt with using latent
variables to achieve unbiased and more efficient results. Researchers
interested in time-invariant predictors can integrate them into a hybrid
FE/RE model with ease. Further extensions, like measurement invariance
testing (Schoot, Lugtig, and Hox
\protect\hyperlink{ref-Schoot2012}{2012}; Millsap
\protect\hyperlink{ref-Millsap2011}{2011}; Steenkamp and Baumgartner
\protect\hyperlink{ref-Steenkamp1998}{1998}) as well as lagged dependent
variables (Bollen and Brand \protect\hyperlink{ref-Bollen2010}{2010};
Allison, Williams, and Moral-Benito
\protect\hyperlink{ref-Allison2017}{2017}) for example, can also be
implemented in a straightforward fashion.

The most basic FE-SEM is furthermore the basis for a variety of
currently popular extended models, such as Latent Curve Models in
general (Curran and Bollen \protect\hyperlink{ref-Curran2001}{2001};
Bollen and Curran \protect\hyperlink{ref-Bollen2004}{2004}), as well as
special implementations like the Dynamic Panel Model (Allison, Williams,
and Moral-Benito \protect\hyperlink{ref-Allison2017}{2017}), the
Random-Intercept Cross-Lagged Panel Model (Hamaker, Kuiper, and Grasman
\protect\hyperlink{ref-Hamaker2015}{2015}) and the Latent Curve Model
with Structured Residuals (Curran et al.
\protect\hyperlink{ref-Curran2014}{2014}). For this reason it is all the
more important for researchers to have a good grasp on the method of
applying FE-SEM. This article is meant to serve as a consolidated
resource for researchers looking for concrete advise on specifying FE
and more general panel models in SEM.

\newpage

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Allison2011}{}%
Allison, Paul. 2011. \emph{Fixed Effects Regression Models}. Thousand
Oaks: Sage Publications.

\leavevmode\hypertarget{ref-Allison2017}{}%
Allison, Paul, Richard Williams, and Enrique Moral-Benito. 2017.
``Maximum Likelihood for Cross-Lagged Panel Models with Fixed Effects.''
\emph{Socius} 3: 1--17. \url{https://doi.org/10.1177/2378023117710578}.

\leavevmode\hypertarget{ref-Bates2010}{}%
Bates, Douglas M. 2010. \emph{Lme4: Mixed-Effects Modeling with R}.
Springer.

\leavevmode\hypertarget{ref-R-lme4}{}%
Bates, Douglas, Martin Mächler, Ben Bolker, and Steve Walker. 2015.
``Fitting Linear Mixed-Effects Models Using lme4.'' \emph{Journal of
Statistical Software} 67 (1): 1--48.
\url{https://doi.org/10.18637/jss.v067.i01}.

\leavevmode\hypertarget{ref-Bell2018}{}%
Bell, Andrew, Malcolm Fairbrother, and Kelvyn Jones. 2018. ``Fixed and
Random Effects Models: Making an Informed Choice.'' \emph{Quality and
Quantity} 53: 1051--74.

\leavevmode\hypertarget{ref-Bollen1989}{}%
Bollen, Kenneth. 1989. \emph{Structural Equations with Latent
Variables}. New York, Chichester: Wiley.

\leavevmode\hypertarget{ref-Bollen2010}{}%
Bollen, Kenneth, and Jennie Brand. 2010. ``A General Panel Model with
Random and Fixed Effects: A Structural Equations Approach.''
\emph{Social Forces} 89(1): 1--34.

\leavevmode\hypertarget{ref-Bollen2004}{}%
Bollen, Kenneth, and Patrick Curran. 2004. ``Autoregressive Latent
Trajectory (ALT) Models: A Synthesis of Two Traditions.''
\emph{Sociological Methods and Research} 32(3): 336--83.
\url{https://doi.org/10.1177/0049124103260222}.

\leavevmode\hypertarget{ref-Bruederl2015}{}%
Brüderl, Josef, and Volker Ludwig. 2015. ``Fixed-Effects Panel
Regression.'' In \emph{The Sage Handbook of Regression Analysis and
Causal Inference}, edited by Henning Best and Christof Wolf, 327--57.
London, Thousand Oaks: Sage Publications.

\leavevmode\hypertarget{ref-Chamberlain1980}{}%
Chamberlain, Gary. 1980. ``Analysis of Covariance with Qualitative
Data.'' \emph{Review of Economic Studies} 47(1): 225--38.

\leavevmode\hypertarget{ref-R-plm_a}{}%
Croissant, Yves, and Giovanni Millo. 2008. ``Panel Data Econometrics in
R: The plm Package.'' \emph{Journal of Statistical Software} 27 (2):
1--43. \url{https://doi.org/10.18637/jss.v027.i02}.

\leavevmode\hypertarget{ref-Curran2011}{}%
Curran, Patrick, and Daniel Bauer. 2011. ``The Disaggregation of
Within-Person and Between-Person Effects in Longitudinal Models of
Change.'' \emph{Annual Review of Psychology} 62: 583--619.

\leavevmode\hypertarget{ref-Curran2001}{}%
Curran, Patrick, and Kenneth Bollen. 2001. ``The Best of Both Worlds:
Combining Autoregressive and Latent Curve Models.'' In \emph{New Methods
for the Analysis of Change}, edited by L. Collins and A. Sayer, 107--35.
Washington, DC: American Psychological Press.
\url{https://doi.org/\%20doi:10.1037/10409-004}.

\leavevmode\hypertarget{ref-Curran2014}{}%
Curran, Patrick, Andrea Howard, Sierra Bainter, Stephanie Lane, and
James McGinley. 2014. ``The Separation of Between-Person and
Within-Person Components of Individual Change over Time: A Latent Growth
Curve Model with Structured Residuals.'' \emph{Journal of Consulting and
Clinical Psychology} 82(5): 879--94.
\url{https://doi.org/10.1037/a0035297}.

\leavevmode\hypertarget{ref-Enders2007}{}%
Enders, Craig, and Davood Tofighi. 2007. ``Centering Predictor Variables
in Cross-Sectional Multilevel Models: A New Look at an Old Issue.''
\emph{Psychological Methods} 12(2): 121--38.

\leavevmode\hypertarget{ref-Graff1979}{}%
Graff, J. 1979. ``Verallgemeinertes LISREL-Modell.'' Mannheim, Germany.

\leavevmode\hypertarget{ref-Hamaker2015}{}%
Hamaker, Ellen, Rebecca Kuiper, and Raoul Grasman. 2015. ``A Critique of
the Cross-Lagged Panel Model.'' \emph{Psychological Methods} 20(1):
102--16. \url{https://doi.org/10.1037/a0038889}.

\leavevmode\hypertarget{ref-Kline2016}{}%
Kline, Rex. 2016. \emph{Principles and Practice of Structural Equation
Modeling. Fourth Edition}. New York: The Guilford Press.

\leavevmode\hypertarget{ref-Millsap2011}{}%
Millsap, Roger. 2011. \emph{Statistical Approaches to Measurement
Invariance}. New York, London: Routledge.

\leavevmode\hypertarget{ref-Mundlak1978}{}%
Mundlak, Yair. 1978. ``On the Pooling of Time Series and Cross Section
Data.'' \emph{Econometrica} 46(1): 69--85.

\leavevmode\hypertarget{ref-Mplus}{}%
Muthén, Linda, and Bengt Muthén. n.d. \emph{Mplus User's Guide. Eighth
Edition}. Los Angeles, California: Muthén \& Muthén.

\leavevmode\hypertarget{ref-R-base}{}%
R Core Team. 2017. \emph{R: A Language and Environment for Statistical
Computing}. Vienna, Austria: R Foundation for Statistical Computing.
\url{https://www.R-project.org/}.

\leavevmode\hypertarget{ref-R-lavaan}{}%
Rosseel, Yves. 2012. ``lavaan: An R Package for Structural Equation
Modeling.'' \emph{Journal of Statistical Software} 48 (2): 1--36.
\url{http://www.jstatsoft.org/v48/i02/}.

\leavevmode\hypertarget{ref-Rosseel2020}{}%
Rosseel, Yves, Terrence D. Jorgensen, Daniel Oberski, Jarrett Byrnes,
Leonard Vanbrabant, Victoria Savalei, Ed Merkle, et al. 2020.
\emph{Lavaan: Latent Variable Analysis}.
\url{https://CRAN.R-project.org/package=lavaan}.

\leavevmode\hypertarget{ref-Schmidheiny2019}{}%
Schmidheiny, Kurt. 2019. ``Panel Data: Fixed and Random Effects.''
Unversität Basel.

\leavevmode\hypertarget{ref-Schoot2012}{}%
Schoot, Rens van de, Peter Lugtig, and Joop Hox. 2012. ``A Checklist for
Testing Measurement Invariance.'' \emph{European Journal of
Developmental Psychology} 9(4): 486--92.

\leavevmode\hypertarget{ref-Schunck2013}{}%
Schunck, Reinhard. 2013. ``Within and Between Estimates in
Random-Effects Models: Advantages and Drawbacks of Random Effects and
Hybrid Models.'' \emph{The Stata Journal} 13(1): 65--76.

\leavevmode\hypertarget{ref-Skrondal2004}{}%
Skrondal, Anders, and Sophia Rabe-Hesketh. 2004. \emph{Generalized
Latent Variable Modeling}. Boca Raton, London, New York, Washington,
D.C.: Chapman \& Hall/CRC.

\leavevmode\hypertarget{ref-Steenkamp1998}{}%
Steenkamp, Jan-Benedict, and Hans Baumgartner. 1998. ``Assessing
Measurement Invariance in Cross-National Consumer Research.''
\emph{Journal of Consumer Research} 25(1): 78--90.

\leavevmode\hypertarget{ref-Teachman2001}{}%
Teachman, Jay, Greg Duncan, Jean Yeung, and Dan Levy. 2001. ``Covariance
Structure Models for Fixed and Random Effects.'' \emph{Sociological
Methods and Research} 30(2): 271--88.

\leavevmode\hypertarget{ref-Wooldridge2002}{}%
Wooldridge, Jeffery. 2002. \emph{Econometric Analysis of Cross Sectional
and Panel Data}. Cambridge, Massachusetts: The MIT Press.

\leavevmode\hypertarget{ref-Wooldridge2012}{}%
---------. 2012. \emph{Introductory Econometrics: A Modern Approach,
5\(^{th}\) Edition}. Mason, Ohio: Thomson South-Western.

\end{document}