### 2020-07-28 article_format/extensions_format.tex

% interactcadsample.tex
% v1.03 - April 2017

\documentclass[]{interact}

\usepackage{epstopdf}% To incorporate .eps illustrations using PDFLaTeX, etc.
\usepackage{subfigure}% Support for small, `sub' figures and tables
%\usepackage[nolists,tablesfirst]{endfloat}% To `separate' figures and tables from text if required

\usepackage{natbib}% Citation support using natbib.sty
\bibpunct[, ]{(}{)}{;}{a}{}{,}% Citation support using natbib.sty
\renewcommand\bibfont{\fontsize{10}{12}\selectfont}% Bibliography support using natbib.sty

\theoremstyle{plain}% Theorem-like structures provided by amsthm.sty
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

\theoremstyle{remark}
\newtheorem{remark}{Remark}
\newtheorem{notation}{Notation}

% see https://stackoverflow.com/a/47122900
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}

\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage{booktabs}
\usepackage{bm}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{tikz}
\usetikzlibrary{arrows}
\usepackage[nofiglist]{endfloat}
\usepackage{blkarray}
\usepackage{setspace}
\usepackage{etoolbox}
\DeclareMathOperator{\E}{\mathbb{E}}
\DeclareMathOperator{\Var}{\mathrm{Var}}
\DeclareMathOperator{\Cov}{\mathrm{Cov}}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\mathtoolsset{showonlyrefs}
\BeforeBeginEnvironment{equation}{\begin{singlespace}\vspace*{-\baselineskip}}
\AfterEndEnvironment{equation}{\end{singlespace}\noindent\ignorespaces}
\BeforeBeginEnvironment{align}{\begin{singlespace}\vspace*{-\baselineskip}}
\AfterEndEnvironment{align}{\end{singlespace}\noindent\ignorespaces}
\def\tightlist{}
\pagenumbering{gobble}

\begin{document}

\articletype{}

\title{Supplementary materials for:\\
A closer look at fixed effects regression in structural equation
modeling using \texttt{lavaan}}

\author{\name{Henrik Kenneth Andersen$^{a}$}
\affil{$^{a}$Chemnitz University of Technology, Institute of Sociology, Chair for
Empirical Social Research, Thüringer Weg 9, 09126 Chemnitz, Germany}
}

\thanks{CONTACT Henrik Kenneth Andersen. Email: \href{mailto:henrik.andersen@soziologie.tu-chemnitz.de}{\nolinkurl{henrik.andersen@soziologie.tu-chemnitz.de}}}

\maketitle

\singlespacing

\doublespacing

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

The following goes into some more detail on the comparison of
traditional FE and FE-SEM models and discusses several opportunities to
extend the basic FE model outlined in the main article, and provides
concrete guidance on the implementation in \texttt{lavaan}. First, I
verify that the FE-SEM model does, in fact, return essentially identical
results compared to the more traditional methods. Then, I go over a
number of possibilities to relax assumptions associated with the
traditional FE model. Then, I discuss the issue of measurement error and
show how we can use latent variables to deal with it and properly
estimate the coefficients of interest. Then, I show a type of hybrid
FE/RE model that allows us to control for time-invariant unobserved
heterogeneity while including time-invariant predictors in the model.

\hypertarget{a-comparison-with-non-sem-methods}{%
\section{A comparison with non-SEM
methods}\label{a-comparison-with-non-sem-methods}}

Just to be sure that the FE-SEM results do, in fact, line up with the
more traditional methods outlined in Section 2 of the main article, we
can use the long-format data (see
\href{https://github.com/henrik-andersen/FE-SEM}{the supplementary
materials at}) to run the typical FE model using the \texttt{plm}
package \citep{R-plm_a}. By default, the \texttt{plm} function assumes
the dataframe is structured so that the first two columns correspond to
the individual and time indices, see the
\href{https://cran.r-project.org/web/packages/plm/plm.pdf}{documentation}
or \citet{R-plm_a}.

\singlespacing

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(plm)}

\CommentTok{# Run the FE model in plm }
\NormalTok{fe1 <-}\StringTok{ }\KeywordTok{plm}\NormalTok{(y }\OperatorTok{~}\StringTok{ }\NormalTok{x, }
           \DataTypeTok{effect =} \StringTok{"individual"}\NormalTok{, }\DataTypeTok{model =} \StringTok{"within"}\NormalTok{, }
           \DataTypeTok{data =}\NormalTok{ df)}
\KeywordTok{summary}\NormalTok{(fe1)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Oneway (individual) effect Within Model
## 
## Call:
## plm(formula = y ~ x, data = df, effect = "individual", model = "within")
## 
## Balanced Panel: n = 1000, T = 5, N = 5000
## 
## Residuals:
##      Min.   1st Qu.    Median   3rd Qu.      Max. 
## -3.720902 -0.601550 -0.021365  0.600833  3.238716 
## 
## Coefficients:
##   Estimate Std. Error t-value  Pr(>|t|)    
## x 0.293907   0.015635  18.798 < 2.2e-16 ***
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Total Sum of Squares:    4452.6
## Residual Sum of Squares: 4091.1
## R-Squared:      0.081192
## Adj. R-Squared: -0.14857
## F-statistic: 353.377 on 1 and 3999 DF, p-value: < 2.22e-16
\end{verbatim}

\doublespacing

From this, we see that the results are, indeed, essentially identical,
with \(\hat{\beta}_{FE-SEM} =0.294 \ (0.016)\) and
\(\hat{\beta}_{FE} = 0.294 \ (0.016)\).

Other methods of estimating FE models work in the random or mixed
effects model framework. For example, we can include the cluster means
per individual of the time-varying independent variables, here \(x\), in
the equation to achieve within estimates
\citep{Mundlak1978, Chamberlain1980, Wooldridge2002}.

\singlespacing

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Generate the cluster means for x per id }
\NormalTok{clusterMeanx <-}\StringTok{ }\KeywordTok{aggregate}\NormalTok{(df}\OperatorTok{$}\NormalTok{x, }\DataTypeTok{by =} \KeywordTok{list}\NormalTok{(df}\OperatorTok{$}\NormalTok{id), }\DataTypeTok{FUN =}\NormalTok{ mean)}
\CommentTok{# Rename the columns }
\KeywordTok{names}\NormalTok{(clusterMeanx) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"id"}\NormalTok{, }\StringTok{"xbar"}\NormalTok{)}

\CommentTok{# Add the cluster means back into df  }
\NormalTok{df <-}\StringTok{ }\KeywordTok{merge}\NormalTok{(df, clusterMeanx, }\DataTypeTok{by =} \StringTok{"id"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\doublespacing

Here using the \texttt{plm} function in the \texttt{random} setup:

\singlespacing

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{fe2 <-}\StringTok{ }\KeywordTok{plm}\NormalTok{(y }\OperatorTok{~}\StringTok{ }\NormalTok{x }\OperatorTok{+}\StringTok{ }\NormalTok{xbar, }
           \DataTypeTok{effect =} \StringTok{"individual"}\NormalTok{, }\DataTypeTok{model =} \StringTok{"random"}\NormalTok{,}
           \DataTypeTok{data =}\NormalTok{ df)}
\KeywordTok{summary}\NormalTok{(fe2)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Oneway (individual) effect Random Effect Model 
##    (Swamy-Arora's transformation)
## 
## Call:
## plm(formula = y ~ x + xbar, data = df, effect = "individual", 
##     model = "random")
## 
## Balanced Panel: n = 1000, T = 5, N = 5000
## 
## Effects:
##                  var std.dev share
## idiosyncratic 1.0230  1.0115 0.862
## individual    0.1637  0.4046 0.138
## theta: 0.2546
## 
## Residuals:
##       Min.    1st Qu.     Median    3rd Qu.       Max. 
## -4.0697324 -0.6786344 -0.0094158  0.6678915  3.5329988 
## 
## Coefficients:
##              Estimate Std. Error z-value Pr(>|z|)    
## (Intercept) -0.022559   0.019192 -1.1754   0.2398    
## x            0.293907   0.015635 18.7983   <2e-16 ***
## xbar         0.757738   0.024015 31.5532   <2e-16 ***
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Total Sum of Squares:    8878.9
## Residual Sum of Squares: 5112.1
## R-Squared:      0.42424
## Adj. R-Squared: 0.42401
## Chisq: 3682.01 on 2 DF, p-value: < 2.22e-16
\end{verbatim}

\doublespacing

And here using the \texttt{lmer} function of the \texttt{lme4} package
\citep{Bates2015} to estimate a mixed model:

\singlespacing

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(lme4)}

\CommentTok{# Run the mixed model in lmer with the cluster means for x }
\NormalTok{mixed1 <-}\StringTok{ }\KeywordTok{lmer}\NormalTok{(y }\OperatorTok{~}\StringTok{ }\NormalTok{x }\OperatorTok{+}\StringTok{ }\NormalTok{xbar }\OperatorTok{+}\StringTok{ }\NormalTok{(}\DecValTok{1} \OperatorTok{|}\StringTok{ }\NormalTok{id), }\DataTypeTok{data =}\NormalTok{ df)}
\KeywordTok{summary}\NormalTok{(mixed1)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Linear mixed model fit by REML ['lmerMod']
## Formula: y ~ x + xbar + (1 | id)
##    Data: df
## 
## REML criterion at convergence: 14906.7
## 
## Scaled residuals: 
##     Min      1Q  Median      3Q     Max 
## -3.9358 -0.6464 -0.0106  0.6349  3.2002 
## 
## Random effects:
##  Groups   Name        Variance Std.Dev.
##  id       (Intercept) 0.1637   0.4046  
##  Residual             1.0230   1.0115  
## Number of obs: 5000, groups:  id, 1000
## 
## Fixed effects:
##             Estimate Std. Error t value
## (Intercept) -0.02256    0.01919  -1.175
## x            0.29391    0.01563  18.798
## xbar         0.75774    0.02401  31.553
## 
## Correlation of Fixed Effects:
##      (Intr) x     
## x     0.000       
## xbar  0.010 -0.651
\end{verbatim}

\doublespacing

In both cases, the models return the same estimates as the FE and FE-SEM
models. Also, in both the \texttt{random} setup using the \texttt{plm}
function, and the mixed model using the \texttt{lmer} function, we get
estimates of the variance components, \(\hat{\sigma}^{2}_{\alpha}\) and
\(\hat{\sigma}^{2}_{\varepsilon}\):

\singlespacing

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Print the variance components for the plm model  }
\KeywordTok{print}\NormalTok{(}\KeywordTok{ercomp}\NormalTok{(fe2), }
      \DataTypeTok{digits =} \DecValTok{3}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##                 var std.dev share
## idiosyncratic 1.023   1.011  0.86
## individual    0.164   0.405  0.14
## theta: 0.255
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Print the variance components for the lmer model }
\KeywordTok{print}\NormalTok{(}\KeywordTok{VarCorr}\NormalTok{(mixed1), }
      \DataTypeTok{comp =} \KeywordTok{c}\NormalTok{(}\StringTok{"Variance"}\NormalTok{, }\StringTok{"Std.Dev"}\NormalTok{), }
      \DataTypeTok{digits =} \DecValTok{3}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  Groups   Name        Variance Std.Dev.
##  id       (Intercept) 0.164    0.405   
##  Residual             1.023    1.011
\end{verbatim}

\doublespacing

From this we see both models report the same estimated variance
components, \(\hat{\sigma}^{2}_{\alpha} = 0.164\) and
\(\hat{\sigma}^{2}_{\varepsilon} = 1.023\), telling us that about 13.8\%
of the residual variance is due to the differences between individuals
(shown in the \texttt{share} column of the \texttt{ercomp()} output).
This is what is referred to as the intraclass correlation coefficient,
or ICC \citep{Hox2010}.

\hypertarget{exten}{%
\section{Extensions}\label{exten}}

\hypertarget{relax}{%
\subsection{Relaxing assumptions meant to mimic traditional FE
models}\label{relax}}

There are a number of implicit assumptions attached to the typical FE
model that can be relaxed in SEM. Some of these assumptions have been
discussed already, and a fairly comprehenisve list of assumptions can be
found in \citet{Bollen2010}. Here, I will go over just a few,
concentrating on the implementation in \texttt{lavaan} and the
opportunity to empirically test whether the adjustments are justified or
not.

The assumptions we will discuss here pertain to the time-invariance of
the effects of both the latent individual effects and the observed
covariates, as well as a time-invariant error variance. We can also
empirically test the correlation between the individual effects and the
covariates to see whether a RE model is preferable to the FE model.

For example, we can rewrite the original FE equation as \begin{align}
y_{it} & = \beta_{t}x_{it} + \lambda_{t}\alpha_{i} + \varepsilon_{it}
\end{align} where \(\beta\) becomes \(\beta_{t}\) and the implicit
regression weight of one turns to \(\lambda_{t}\) to highlight the fact
that the effect of \(x\) as well as \(\alpha\) on \(y\) may vary over
time. We can furthermore easily relax the assumption of time-constant
error variance, i.e., \(\sigma^{2}_{\varepsilon_{t}}\). As noted in the
\href{https://github.com/henrik-andersen/FE-SEM/blob/master/article.pdf}{main
article}, the assumption regarding
\(\mathop{\mathrm{\mathbb{E}}}[\alpha x_{t}]\) in \(\bm{\Psi}\)
determines whether we have an FE or RE model. We can set these to zero
and test whether the RE model would be preferable to the FE model. In
general, if the individual effects are truly uncorrelated with the model
covariates, it is advisable to switch to an RE model since because it
uses up less degrees of freedom, it will have smaller standard errors
\citep{Bollen2010}.

In the following \texttt{lavaan} code, we simply remove the factor
loadings of one for the latent individual effect variable which allows
them to be estimated freely at each timepoint. For the effect of the
covariate, we can either delete the constraints \texttt{b} in
\texttt{yt\ \textasciitilde{}\ b*xt} or give each regression a different
label, e.g., \texttt{b1}, \texttt{b2}, \texttt{b3}, etc. Similarly, to
allow the error variance to vary over time, we turn the constraints
\texttt{e} into simple labels, i.e., \texttt{e1}, \texttt{e2},
\texttt{e3}, etc., or again just delete them. In fact, regarding the
error variances, they will be estimated necessarily, and do not need to
be explicitly mentioned in the model syntax at all. Finally, to move
from an FE to an RE model, we could simply constrain the correlations
between the individual effects and the covariates to zero, i.e.,
\texttt{a\ \textasciitilde{}\textasciitilde{}\ 0*x1\ +\ 0*x2\ +\ 0*x3\ +\ 0*x4\ +\ 0*x5}.

\singlespacing

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{fe_sem_fullyrelaxed <-}\StringTok{ '}
\StringTok{# Define individual effects variable }
\StringTok{a =~ y1 + y2 + y3 + y4 + y5}
\StringTok{# Regressions, constrain coefficient to be equal over time}
\StringTok{y1 ~ b1*x1}
\StringTok{y2 ~ b2*x2 }
\StringTok{y3 ~ b3*x3}
\StringTok{y4 ~ b4*x4}
\StringTok{y5 ~ b5*x5}
\StringTok{# Allow unrestricted correlation between eta and covariates}
\StringTok{a ~~ x1 + x2 + x3 + x4 + x5}
\StringTok{# Alternatively: constrain all to 0 for RE model, or}
\StringTok{# just individual correlations}
\StringTok{# a ~~ 0*x1 + 0*x2 + 0*x3 + 0*x4 + 0*x5}
\StringTok{x1 ~~ x2 + x3 + x4 + x5}
\StringTok{x2 ~~ x3 + x4 + x5}
\StringTok{x3 ~~ x4 + x5}
\StringTok{x4 ~~ x5}
\StringTok{# Constrain residual variances to be equal over time}
\StringTok{y1 ~~ e1*y1}
\StringTok{y2 ~~ e2*y2}
\StringTok{y3 ~~ e3*y3}
\StringTok{y4 ~~ e4*y4}
\StringTok{y5 ~~ e5*y5}
\StringTok{'}
\NormalTok{fe_sem_fullyrelaxed.fit <-}\StringTok{ }\KeywordTok{sem}\NormalTok{( }\DataTypeTok{model =}\NormalTok{ fe_sem_fullyrelaxed, }
                                \DataTypeTok{data =}\NormalTok{ dfw, }
                                \DataTypeTok{estimator =} \StringTok{"ML"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\doublespacing

\singlespacing

\doublespacing

As outlined in \citet{Bollen2010}, the researcher has the opportunity to
test each of the assumptions empirically and decide whether a more
parsimonious, i.e., restrictive model is justifiable. For each
assumption, a likelihood ratio test can be carried out to determine
whether the improvement to model fit resulting from the relaxation of
various assumptions is significant or whether the more parsimonious
model is preferable after all.

If we use the original model \texttt{fe\_sem.fit} (from the
\href{https://github.com/henrik-andersen/FE-SEM/blob/master/article.pdf}{main
article}) as a starting point, the best strategy for testing these
assumptions is to work in a stepwise fashion, relaxing one assumption at
a time. We can begin by first constraining the correlation between
\(\alpha\) and \(x_{t}\) to zero (\texttt{re\_sem}) for an RE model. If
turning from an FE to an RE model does not significantly worsen model
fit, we can go forward with the rest of the steps with the RE model. If,
however, the fit does worsen significantly, it is likely better to stick
with the FE model; moving forward then with it to see if a less
restrictive FE model is preferable. We can perform a likelihood ratio
test in \texttt{R} using the \texttt{anova()} function:

\small
\singlespacing

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{anova}\NormalTok{( fe_sem.fit, re_sem.fit)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Chi-Squared Difference Test
## 
##            Df   AIC   BIC   Chisq Chisq diff Df diff Pr(>Chisq)    
## fe_sem.fit 32 30998 31111  30.137                                  
## re_sem.fit 37 31809 31897 850.928     820.79       5  < 2.2e-16 ***
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
\end{verbatim}

\doublespacing
\normalsize

The table that is generated shows a comparison of the nested models, in
decending order according to degrees of freedom. The RE model does not
estimate the correlations between the individual effects and the
covariates, so it is more parsimonious and thus listed at the bottom.
The \texttt{Chisq} column shows the \(\chi^{2}\) statistic for both
models and the \texttt{Chisq\ diff} column calculates the difference
between the two. Obviously, according to the DGP, the correlation
between the individual effects and \(x_{t}\) is not zero, so fixing
these to zero leads to a substantial amount of misfit. The last column
puts the \(\chi^{2}\) difference in relation to the difference in
degrees of freedom and gives a p-value for the probability that the
difference is solely due to chance. Here, the change in \(\chi^{2}\) is
highly significant, so the FE model should be retained.

After now having established once and for all that FE is our preferred
model, we can begin relaxing the rest of the assumptions. I show the
following merely as a demonstration of the procedure, we know already
from the DGP that the parsimonious model as specified in
\texttt{fe\_sem.fit} is appropriate. We can next allow the error
variances (\texttt{fe\_semb.fit}), the effect of \(x\) on \(y\)
(\texttt{fe\_semc.fit}) and finally the factor loadings of the
individual effects (\texttt{fe\_semd.fit}) all to vary over time.

\singlespacing

\doublespacing

\small
\singlespacing

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{anova}\NormalTok{( fe_sem.fit, fe_semb.fit, fe_semc.fit, fe_semd.fit)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Chi-Squared Difference Test
## 
##             Df   AIC   BIC  Chisq Chisq diff Df diff Pr(>Chisq)
## fe_semd.fit 20 31017 31189 25.140                              
## fe_semc.fit 24 31010 31162 25.764     0.6249       4     0.9603
## fe_semb.fit 28 31003 31135 26.686     0.9215       4     0.9215
## fe_sem.fit  32 30998 31111 30.137     3.4516       4     0.4853
\end{verbatim}

\doublespacing
\normalsize

Keep in mind that a less parsimonious model (fewer degrees of freedom)
can never fit worse than a more parsimonious one (more degrees of
freedom). I.e., chance variations due to sampling error mean that adding
constraints to a model will tend to always worsen fit, at least
minimally. The question here is whether the improvement to fit by
loosening constraints is meaningful or not. In the table above, we
should not expect any meaningful improvements moving from
\texttt{fe\_sem.fit} to \texttt{fe\_semd.fit}. Here, using simulated
data, we have the luxury of knowing that any significant differences in
\(\chi^{2}\) are due to chance. With real data, it is up to the
researcher to apply their best judgment and decide whether the results
are plausible or not.

\hypertarget{measerr}{%
\subsection{Measurement error}\label{measerr}}

What if the observed variables are not measured perfectly? Then what we
observe, call them \(\tilde{x}_{t}\) and \(\tilde{y}_{t}\) are
composites of the true score we are after, i.e., \(x_{t}\) and
\(y_{t}\), plus an additive measurement error portion: \begin{align}
\tilde{x}_{t} & = x_{t} + \upsilon_{t}, \\
\tilde{y}_{t} & = y_{t} + \nu_{t}.
\end{align} How does this affect our model? Well, first notice that
measurement error in the dependent variable is typically less of a
serious problem than measurement error in the independent variables. Let
us assume again mean-centered variables so that we can ignore the
intercept, and consider the following simple bivariate equation:
\begin{align}
y & = \beta x + \varepsilon
\end{align} if \(y\) is measured imperfectly and what we observe is
\(\tilde{y} = y + \nu\), then we can rewrite the equation as:
\begin{align}
(\tilde{y} - \nu) & = \beta x + \varepsilon \\
\tilde{y} & = \beta x + \varepsilon + \nu.
\end{align} The measurement error in \(y\) just gets added to the
regression error. As long as \(\nu\) is uncorrelated with \(x\), then
the regression coefficient will be unbiased
\citep{Pischke2007, Wooldridge2009}. However, this will increase the
error variance and thus make the estimates less precise.

We will look at the effect of measurement error in the dependent
variable using an example shortly. For now though, let us be safe in the
knowledge that the coefficient of interest is likely unbiased, and
concentrate on the more serious problem of error in the independent
variable.

The intuition behind the problem of measurement error in the independent
variable(s) can be explained as follows. Take
\(\tilde{x} = x + \upsilon\) and substitute this into the equation for
\(y\): \begin{align}
y & = \beta x + \varepsilon \\
 & = \beta(\tilde{x} - \upsilon) + \varepsilon \\
 & = \beta\tilde{x} + (\varepsilon - \beta\upsilon).
\end{align} Since \(\tilde{x}\) is obviously correlated with
\(\upsilon\) (unless the variance of \(\upsilon\) is so small so that
the correlation is essentially negligible), then the composite error in
this regression is also correlated with the independent variable and
thus the estimated coefficient of \(\beta\) will be biased.

\hypertarget{the-consequences-of-measurement-error}{%
\subsubsection{The consequences of measurement
error}\label{the-consequences-of-measurement-error}}

To demonstrate the effect of measurement error on the FE-SEM model, and
then provide a strategy for dealing with measurement error in SEM, the
\href{https://github.com/henrik-andersen/FE-SEM/blob/master/simulation-code.R}{simulated
dataset} generates multiple \emph{indicators} of the independent and
dependent variables that all measure the intended variable imprecisely.
Returning to our panel data, we have three indicators of each the
independent and dependent variable, per timepoint:\\
\begin{align}
\tilde{x}_{kt} & = x_{t} + \upsilon_{kt}, \\
\tilde{y}_{kt} & = y_{t} + \nu_{kt}
\end{align} where \(k = 1, 2, 3\) and \(t = 1, ..., T\). This is like
repeatedly presenting a respondent with a multi-item scale designed to
measure things like stress, depression, xenophobia, etc. over the course
of a panel study. To create the observed indicators, a random amount of
measurement error (ranging from
\(\{\sigma^{2}_{\upsilon_{k}}, \sigma^{2}_{\nu_{k}}\}\in \{1.0, 1.1, 1.2, 1.3, 1.4, 1.5\}\))
was added to the true variables, again see
\href{https://github.com/henrik-andersen/FE-SEM/blob/master/simulation-code.R}{the
simulation code}.

Let us first focus on the issue of imprecise measurements of the
independent variable of interest and run the same FE-SEM model above,
but this time we will use one of the measurement error sullied
indicators, here \(\tilde{x}_{1t}\), instead of the true independent
variable, \(x_{t}\). As for the naming conventions in the \texttt{R}
code, \texttt{x11} stands for the first indicator (\(k = 1\)) at the
first point in time (\(t = 1\)), whereas for example \texttt{x35} stands
for the third indicator (\(k = 3\)) at the fifth point in time
(\(t = 5\)).

\singlespacing

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{fe_sem2 <-}\StringTok{ '}
\StringTok{# Define individual effects variable }
\StringTok{a =~ 1*y1 + 1*y2 + 1*y3 + 1*y4 + 1*y5}
\StringTok{# Regressions, constrain coefficient to be equal over time}
\StringTok{# Now the imprecisely measured indicator tilde\{x\}_kt}
\StringTok{# instead of the true variable x_t}
\StringTok{y1 ~ b*x11 }
\StringTok{y2 ~ b*x12 }
\StringTok{y3 ~ b*x13}
\StringTok{y4 ~ b*x14}
\StringTok{y5 ~ b*x15}
\StringTok{# Allow unrestricted correlation between eta and covariates}
\StringTok{a ~~ x11 + x12 + x13 + x14 + x15}
\StringTok{x11 ~~ x12 + x13 + x14 + x15}
\StringTok{x12 ~~ x13 + x14 + x15}
\StringTok{x13 ~~ x14 + x15}
\StringTok{x14 ~~ x15}
\StringTok{# Constrain residual variances to be equal over time}
\StringTok{y1 ~~ e*y1}
\StringTok{y2 ~~ e*y2}
\StringTok{y3 ~~ e*y3}
\StringTok{y4 ~~ e*y4}
\StringTok{y5 ~~ e*y5}
\StringTok{'}
\NormalTok{fe_sem2.fit <-}\StringTok{ }\KeywordTok{sem}\NormalTok{( }\DataTypeTok{model =}\NormalTok{ fe_sem2, }
                    \DataTypeTok{data =}\NormalTok{ dfw, }
                    \DataTypeTok{estimator =} \StringTok{"ML"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\doublespacing

Now, for the sake of brevity, let us look just at the estimated
coefficients for \(\beta\).

\singlespacing

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{summary}\NormalTok{( fe_sem2.fit)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
...
##                    Estimate  Std.Err  z-value  P(>|z|)
##   y1 ~                                                
##     x11        (b)    0.091    0.009    9.611    0.000
##   y2 ~                                                
##     x12        (b)    0.091    0.009    9.611    0.000
##   y3 ~                                                
##     x13        (b)    0.091    0.009    9.611    0.000
##   y4 ~                                                
##     x14        (b)    0.091    0.009    9.611    0.000
##   y5 ~                                                
##     x15        (b)    0.091    0.009    9.611    0.000
## 
...
\end{verbatim}

\doublespacing

Obviously, the estimated coefficient \(\hat{\beta} =\) 0.091 is
substantially smaller than the true population coefficient of
\(\beta = 0.3\). And the discrepancy is not just due to sampling error.
In fact, we can derive the bias we are observing here.

For a simple bivariate regression model, it is straightforward to
quantify the bias due to measurement error. It will be \begin{align}
\mathop{\mathrm{\mathrm{Cov}}}(y, \tilde{x}) & = \mathop{\mathrm{\mathbb{E}}}[y \tilde{x}] \\
 & = \mathop{\mathrm{\mathbb{E}}}[(\beta\tilde{x} + \varepsilon)\tilde{x}] \\
 & = \mathop{\mathrm{\mathbb{E}}}[\beta\tilde{x}^{2} + \varepsilon \tilde{x}] \\
 & = \beta \mathop{\mathrm{\mathrm{Var}}}(\tilde{x}) \\
\hat{\beta} & = \frac{\mathop{\mathrm{\mathrm{Cov}}}(y, \tilde{x})}{\mathop{\mathrm{\mathrm{Var}}}(\tilde{x})} \\
 & = \frac{\mathop{\mathrm{\mathbb{E}}}[(\beta x + \varepsilon)(x + \upsilon)]}{\mathop{\mathrm{\mathbb{E}}}[(x + \upsilon)^{2}]} \\
 & = \frac{\mathop{\mathrm{\mathbb{E}}}[\beta x^{2} + \beta x \upsilon + \varepsilon x + \varepsilon \upsilon]}{\mathop{\mathrm{\mathbb{E}}}[x^{2} + 2 x \upsilon + \upsilon^{2}]} \\
 & = \beta \frac{\mathop{\mathrm{\mathrm{Var}}}(x)}{\mathop{\mathrm{\mathrm{Var}}}(x) + \mathop{\mathrm{\mathrm{Var}}}(\upsilon)}.
\end{align} which results if we assume that
\(\mathop{\mathrm{\mathbb{E}}}[x \upsilon] = 0\),
\(\mathop{\mathrm{\mathbb{E}}}[x \varepsilon] = 0\),
\(\mathop{\mathrm{\mathbb{E}}}[\tilde{x} \varepsilon] = 0\) and
\(\mathop{\mathrm{\mathbb{E}}}[\varepsilon \upsilon] = 0\)
\citep{Wooldridge2009}. However, the model we are interested is not a
bivariate model, so what was the point of showing the this? For one, it
points out that the bias will always move the estimated coefficient
closer to 0, since
\(\mathop{\mathrm{\mathrm{Var}}}(x) \le \mathop{\mathrm{\mathrm{Var}}}(x) + \mathop{\mathrm{\mathrm{Var}}}(\upsilon)\).
This means positive effects will be biased downwards and negative
effects biased upwards, always towards zero. This is why it is referred
to as \emph{attenuation bias}. Second, it will help to familiarize
ourselves with this equation to better understand the one for the
multivariate case.

Indeed, the magnitude of the bias in a multivariate model is somewhat
more complex to derive, but it will be \begin{align}
\hat{\beta} & = \beta \frac{\mathop{\mathrm{\mathrm{Var}}}(\theta)}{\mathop{\mathrm{\mathrm{Var}}}(\theta) + \mathop{\mathrm{\mathrm{Var}}}(\upsilon)}
\end{align} where \(\theta\) is just the residual of a regression in
which the true underlying variable is regressed on all other covariates.
In this case, we need to regress \(x_{t}\) on \(\alpha_{1}\) and
\(\alpha_{2}\) for:
\(x_{t} = \tau + \gamma_{1}\alpha_{1} + \gamma_{2}\alpha_{2} + \theta_{t}\)
where \(\tau\) is the intercept, and \(\gamma_{1}\), \(\gamma_{2}\) are
the regression coefficients and \(\theta_{t}\) is the residual
\citep[p.~318--320]{Wooldridge2009}.

Normally it is not possible to reconstruct the bias since in cases where
we have to rely on indicators, we would not know the true underlying
variable. Furthermore, in the case of a fixed-effects model, the
covariates are the unobserved time-invariant characteristics. However,
because we are working with simulated data, we have everything we need.
Going back to the results above, we can get the residuals of \(x_{t}\)
by either running a regression and saving the residuals, or we could
skip a step and get them directly using the `residual maker' matrix
\citep{Ruettenauer2020} which is
\(\bm{M} = \bm{I} - \bm{A}(\bm{A}^{\intercal}\bm{A})^{-1}\bm{A}^{\intercal}\)
and
\(\bm{A} = \begin{pmatrix}\bm{\iota}_{n} & \bm{\alpha_{1}} & \bm{\alpha_{2}}\end{pmatrix}\)
is the \(n \times 3\) matrix of the covariates (plus a constant).

\singlespacing

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Make the n x n identity matrix}
\NormalTok{Id <-}\StringTok{ }\KeywordTok{diag}\NormalTok{( n)}

\CommentTok{# n x 2 matrix of covariates a1 and a2}
\NormalTok{A <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{( }\KeywordTok{c}\NormalTok{( }\KeywordTok{rep}\NormalTok{( }\DecValTok{1}\NormalTok{, n), dfw}\OperatorTok{$}\NormalTok{a1, dfw}\OperatorTok{$}\NormalTok{a2), }
             \DataTypeTok{nrow =}\NormalTok{ n, }\DataTypeTok{ncol =} \DecValTok{3}\NormalTok{)}

\CommentTok{# The residual maker matrix M = I - A(A'A)^-1 A'}
\NormalTok{M <-}\StringTok{ }\NormalTok{Id }\OperatorTok{-}\StringTok{ }\NormalTok{A }\OperatorTok{%*%}\StringTok{ }\KeywordTok{solve}\NormalTok{( }\KeywordTok{t}\NormalTok{( A) }\OperatorTok{%*%}\StringTok{ }\NormalTok{A) }\OperatorTok{%*%}\StringTok{ }\KeywordTok{t}\NormalTok{( A)}

\CommentTok{# Save the residuals, t for 'theta'}
\NormalTok{t <-}\StringTok{ }\NormalTok{M }\OperatorTok{%*%}\StringTok{ }\NormalTok{dfw}\OperatorTok{$}\NormalTok{x1}

\CommentTok{# Re-run the FE model from above with the 'true' }
\CommentTok{# independent variable for the correct estimate for beta }
\NormalTok{fe_sem.fit <-}\StringTok{ }\KeywordTok{sem}\NormalTok{( }\DataTypeTok{model =}\NormalTok{ fe_sem, }\DataTypeTok{data =}\NormalTok{ dfw, }\DataTypeTok{estimator =} \StringTok{"ML"}\NormalTok{)}

\CommentTok{# The equation for the biased beta }
\KeywordTok{lavInspect}\NormalTok{( fe_sem.fit, }\StringTok{"list"}\NormalTok{)[ }\DecValTok{6}\NormalTok{, }\DecValTok{14}\NormalTok{]}\OperatorTok{*}
\StringTok{  }\NormalTok{(( }\KeywordTok{var}\NormalTok{( t))}\OperatorTok{/}\NormalTok{( }\KeywordTok{var}\NormalTok{( t) }\OperatorTok{+}\StringTok{ }\KeywordTok{var}\NormalTok{( dfw}\OperatorTok{$}\NormalTok{x11 }\OperatorTok{-}\StringTok{ }\NormalTok{dfw}\OperatorTok{$}\NormalTok{x1)))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##           [,1]
## [1,] 0.1027325
\end{verbatim}

\doublespacing

\singlespacing

\doublespacing

From this we can see that the biased estimate above of \(\hat{\beta} =\)
0.091 roughly comes from
\(\beta \frac{\mathop{\mathrm{\mathrm{Var}}}(\theta_{t})}{\mathop{\mathrm{\mathrm{Var}}}(\theta_{t}) + \mathop{\mathrm{\mathrm{Var}}}(\upsilon_{t})} = 0.294 \frac{1.054}{3.018}\)
= 0.103; `roughly' because the equation here is the population equation.
Due to sampling error, the estimates will tend vary slightly.

\hypertarget{using-latent-variables-to-deal-with-measurement-error}{%
\subsubsection{Using latent variables to deal with measurement
error}\label{using-latent-variables-to-deal-with-measurement-error}}

The way we deal with measurement error in SEM is surprisingly similar to
the logic of fixed-effects regression. Namely, if we have multiple
cross-sectional observations of the underlying construct of interest,
then we can define a latent variable that represents the common variance
across those multiple variables. Contrast this with the use of
longitudinal repeated measures to isolate the common variance across
time.

So, if we do in fact have multiple cross-sectional indicators for the
underlying variables of interest, then we can partition them into an
explained and unexplained portion:\\
\begin{align}
x_{kt} & = \lambda_{kt}^{x}\xi_{t} + \delta_{kt}, \\
y_{kt} & = \lambda_{kt}^{y}\eta_{t} + \varepsilon_{kt},
\end{align} where \(x_{kt}\) and \(y_{kt}\) are the \(k^{th}\)
indicators, \(\xi_{t}\) and \(\eta_{t}\) are latent factors representing
the common variance across the cross-sectional repeated measures, and
\(\delta_{kt}\) and \(\varepsilon_{kt}\) are the unexplained portions of
\(x_{t}\) and \(y_{t}\), respectively. The latent factors are linked to
the observed indicators through the factor loadings \(\lambda_{kt}\).

Thus, our FE regression equation changes from
\(y_{t} = \beta x_{t} + \alpha + \varepsilon_{t}\) to: \begin{align}
\eta_{t} & = \beta \xi_{t} + \alpha + \zeta_{t}
\end{align} where \(\zeta_{t}\) represents the disturbance, in other
words the residual of the latent dependent variable \(\eta_{t}\). First,
however, let us double-check that measurement error in the dependent
variable only increases the error variance (thus also increasing
standard errors and reducing \(R^{2}\)), but does not systematically
bias the coefficients of interest. The next model uses the indicators of
\(x\) and specifies latent variables (\(\xi_{t}\), \texttt{xi} in the
code) to represent the valid cross-sectional variance. The dependent
variable in the model is one of the impercisely measured indicators of
\(y\).

\singlespacing

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{fe_sem3 <-}\StringTok{ '}
\StringTok{# Define individual effects variable }
\StringTok{a =~ 1*y11 + 1*y12 + 1*y13 + 1*y14 + 1*y15}
\StringTok{# Measurement model for independent variables, xi }
\StringTok{xi1 =~ 1*x11 + x21 + x31 }
\StringTok{xi2 =~ 1*x12 + x22 + x32}
\StringTok{xi3 =~ 1*x13 + x23 + x33}
\StringTok{xi4 =~ 1*x14 + x24 + x34}
\StringTok{xi5 =~ 1*x15 + x25 + x35}
\StringTok{# Regressions, constrain coefficient to be equal over time}
\StringTok{y11 ~ b*xi1}
\StringTok{y12 ~ b*xi2 }
\StringTok{y13 ~ b*xi3}
\StringTok{y14 ~ b*xi4}
\StringTok{y15 ~ b*xi5}
\StringTok{# Allow unrestricted correlation between eta and covariates}
\StringTok{a ~~ xi1 + xi2 + xi3 + xi4 + xi5}
\StringTok{xi1 ~~ xi2 + xi3 + xi4 + xi5}
\StringTok{xi2 ~~ xi3 + xi4 + xi5}
\StringTok{xi3 ~~ xi4 + xi5}
\StringTok{xi4 ~~ xi5}
\StringTok{# Constrain residual variances to be equal over time}
\StringTok{y11 ~~ e*y11}
\StringTok{y12 ~~ e*y12}
\StringTok{y13 ~~ e*y13}
\StringTok{y14 ~~ e*y14}
\StringTok{y15 ~~ e*y15}
\StringTok{'}
\NormalTok{fe_sem3.fit <-}\StringTok{ }\KeywordTok{sem}\NormalTok{( }\DataTypeTok{model =}\NormalTok{ fe_sem3, }
                    \DataTypeTok{data =}\NormalTok{ dfw, }
                    \DataTypeTok{estimator =} \StringTok{"ML"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\doublespacing

\singlespacing

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{summary}\NormalTok{( fe_sem3.fit)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
...
## Regressions:
##                    Estimate  Std.Err  z-value  P(>|z|)
##   y11 ~                                               
##     xi1        (b)    0.299    0.029   10.302    0.000
##   y12 ~                                               
##     xi2        (b)    0.299    0.029   10.302    0.000
##   y13 ~                                               
##     xi3        (b)    0.299    0.029   10.302    0.000
##   y14 ~                                               
##     xi4        (b)    0.299    0.029   10.302    0.000
##   y15 ~                                               
##     xi5        (b)    0.299    0.029   10.302    0.000
...
\end{verbatim}

\doublespacing

The estimated coefficient here in model \texttt{fe\_sem3.fit} is
\(\hat{\beta}_{y_{1t},\xi_{t}} =\) 0.299 which is very close to the
estimated coefficient in the first, correctly specified model
\texttt{fe\_sem.fit}, where \(\hat{\beta}_{y_{t},x_{t}} =\) 0.294.
Notice, however, that the standard error of the estimate is
substantially larger, with 0.029 in \texttt{fe\_sem3.fit} vs.~0.016 in
\texttt{fe\_sem.fit} in which \(y\) was measured without error. The
explained variance (\(R^{2}\)) in the dependent variable was also much
higher in the first model:

\singlespacing

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{lavInspect}\NormalTok{( fe_sem.fit, }\StringTok{"r2"}\NormalTok{)[ }\DecValTok{1}\OperatorTok{:}\DecValTok{5}\NormalTok{]}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##        y1        y2        y3        y4        y5 
## 0.5893174 0.5928913 0.5894869 0.5853815 0.5845328
\end{verbatim}

\doublespacing

compared to the current model:

\singlespacing

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{lavInspect}\NormalTok{( fe_sem3.fit, }\StringTok{"r2"}\NormalTok{)[ }\DecValTok{1}\OperatorTok{:}\DecValTok{5}\NormalTok{]}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##       y11       y12       y13       y14       y15 
## 0.3901187 0.3914316 0.3828739 0.3850362 0.3703136
\end{verbatim}

\doublespacing

Finally, to see the benefits of removing measurement error from the
dependent variable in terms of standard errors and \(R^{2}\) statistics,
we can specify a model with latent variables representing the valid
cross-sectional variance in \(y\) (\texttt{n} for \(\eta\) in the code).

\singlespacing

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{fe_sem4 <-}\StringTok{ '}
\StringTok{# Measurement model for dependent variable, n for eta}
\StringTok{n1 =~ 1*y11 + y21 + y31}
\StringTok{n2 =~ 1*y12 + y22 + y32}
\StringTok{n3 =~ 1*y13 + y23 + y33}
\StringTok{n4 =~ 1*y14 + y24 + y34}
\StringTok{n5 =~ 1*y15 + y25 + y35}
\StringTok{# Define individual effects variable }
\StringTok{a =~ 1*n1 + 1*n2 + 1*n3 + 1*n4 + 1*n5}
\StringTok{# Measurement model for independent variables, xi }
\StringTok{xi1 =~ 1*x11 + x21 + x31 }
\StringTok{xi2 =~ 1*x12 + x22 + x32}
\StringTok{xi3 =~ 1*x13 + x23 + x33}
\StringTok{xi4 =~ 1*x14 + x24 + x34}
\StringTok{xi5 =~ 1*x15 + x25 + x35}
\StringTok{# Regressions, constrain coefficient to be equal over time}
\StringTok{n1 ~ b*xi1}
\StringTok{n2 ~ b*xi2 }
\StringTok{n3 ~ b*xi3}
\StringTok{n4 ~ b*xi4}
\StringTok{n5 ~ b*xi5}
\StringTok{# Allow unrestricted correlation between eta and covariates}
\StringTok{a ~~ xi1 + xi2 + xi3 + xi4 + xi5}
\StringTok{xi1 ~~ xi2 + xi3 + xi4 + xi5}
\StringTok{xi2 ~~ xi3 + xi4 + xi5}
\StringTok{xi3 ~~ xi4 + xi5}
\StringTok{xi4 ~~ xi5}
\StringTok{# Constrain residual variances to be equal over time}
\StringTok{n1 ~~ e*n1}
\StringTok{n2 ~~ e*n2}
\StringTok{n3 ~~ e*n3}
\StringTok{n4 ~~ e*n4}
\StringTok{n5 ~~ e*n5}
\StringTok{'}
\NormalTok{fe_sem4.fit <-}\StringTok{ }\KeywordTok{sem}\NormalTok{( }\DataTypeTok{model =}\NormalTok{ fe_sem4, }
                    \DataTypeTok{data =}\NormalTok{ dfw, }
                    \DataTypeTok{estimator =} \StringTok{"ML"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\doublespacing

\singlespacing

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{summary}\NormalTok{( fe_sem4.fit)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
...
## Regressions:
##                    Estimate  Std.Err  z-value  P(>|z|)
##   n1 ~                                                
##     xi1        (b)    0.264    0.023   11.515    0.000
##   n2 ~                                                
##     xi2        (b)    0.264    0.023   11.515    0.000
##   n3 ~                                                
##     xi3        (b)    0.264    0.023   11.515    0.000
##   n4 ~                                                
##     xi4        (b)    0.264    0.023   11.515    0.000
##   n5 ~                                                
##     xi5        (b)    0.264    0.023   11.515    0.000
...
\end{verbatim}

\doublespacing

Here, the effect \(\hat{\beta}_{\eta_{t},\xi_{t}}\) is somewhat further
off of the true effect of 0.3 than the preceding models. This will
depend on how the latent variables are estimated, which themselves will
depend on the underlying correlations between the indicators. Again, if
the main goal of the model is to avoid bias, it may be advisable to just
leave the manifest dependent variable as it is, and worry about
measurement error in the independent variables.

\hypertarget{time-invariant-predictors}{%
\subsection{Time-invariant predictors}\label{time-invariant-predictors}}

What if we do not just want to just control for the effects of all
time-invariant variables, but investigate some of them in detail? Many
time-invariant variables, like sex, birth cohort, nationality,
education, etc. can be interesting on their own. And typically, many of
these variables are readily available in a given dataset. The
traditional OLS-based FE model does not allow for this, as it wipes out
the effect of \emph{all} time-invariant variables, whether observed or
not.

In SEM, we can easily specify a type of \emph{hybrid} FE/RE model
\citep{Bollen2010} that allows us to control for time-invariant
unobserved heterogeneity while also investigating the effects of
specific observed time-invariant predictors.\footnote{These types of
  models have become well known outside of SEM as well, see for example
  \citet{Allison2011}; \citet{Schunck2013}; \citet{Bell2018}.}

In the next example, we continue with the most complex model we have
specified so far, \texttt{fe\_sem4.fit} in which measurement error in
both the independent and dependent variables is accounted for using
latent variables. Now, we would like as well to specifically investigate
the effect of \(\alpha_{2}\) on the dependent variable. The equation for
this model changes to:
\(\eta_{t} = \beta \xi_{t} + \alpha + \gamma \alpha_{2} + \zeta_{t}\).

\singlespacing

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{fe_sem5 <-}\StringTok{ '}
\StringTok{# Measurement model for dependent variable, n for eta}
\StringTok{n1 =~ 1*y11 + y21 + y31}
\StringTok{n2 =~ 1*y12 + y22 + y32}
\StringTok{n3 =~ 1*y13 + y23 + y33}
\StringTok{n4 =~ 1*y14 + y24 + y34}
\StringTok{n5 =~ 1*y15 + y25 + y35}
\StringTok{# Define individual effects variable }
\StringTok{a =~ 1*n1 + 1*n2 + 1*n3 + 1*n4 + 1*n5}
\StringTok{# Measurement model for independent variables, xi }
\StringTok{xi1 =~ 1*x11 + x21 + x31 }
\StringTok{xi2 =~ 1*x12 + x22 + x32}
\StringTok{xi3 =~ 1*x13 + x23 + x33}
\StringTok{xi4 =~ 1*x14 + x24 + x34}
\StringTok{xi5 =~ 1*x15 + x25 + x35}
\StringTok{# Regressions, constrain coefficient to be equal over time}
\StringTok{n1 ~ b*xi1 + g*a2}
\StringTok{n2 ~ b*xi2 + g*a2}
\StringTok{n3 ~ b*xi3 + g*a2}
\StringTok{n4 ~ b*xi4 + g*a2}
\StringTok{n5 ~ b*xi5 + g*a2}
\StringTok{# Allow unrestricted correlation between eta and covariates}
\StringTok{a ~~ xi1 + xi2 + xi3 + xi4 + xi5 + 0*a2}
\StringTok{a2 ~~ xi1 + xi2 + xi3 + xi4 + xi5}
\StringTok{xi1 ~~ xi2 + xi3 + xi4 + xi5}
\StringTok{xi2 ~~ xi3 + xi4 + xi5}
\StringTok{xi3 ~~ xi4 + xi5}
\StringTok{xi4 ~~ xi5}
\StringTok{# Constrain residual variances to be equal over time}
\StringTok{n1 ~~ e*n1}
\StringTok{n2 ~~ e*n2}
\StringTok{n3 ~~ e*n3}
\StringTok{n4 ~~ e*n4}
\StringTok{n5 ~~ e*n5}
\StringTok{'}
\NormalTok{fe_sem5.fit <-}\StringTok{ }\KeywordTok{sem}\NormalTok{( }\DataTypeTok{model =}\NormalTok{ fe_sem5, }
                    \DataTypeTok{data =}\NormalTok{ dfw, }
                    \DataTypeTok{estimator =} \StringTok{"ML"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\doublespacing

Keep in mind, based on the DGP, the true parameters are \(\beta = 0.3\)
and \(\gamma = 0.45\).

\singlespacing

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{summary}\NormalTok{( fe_sem5.fit)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
...
## Regressions:
##                    Estimate  Std.Err  z-value  P(>|z|)
##   n1 ~                                                
##     xi1        (b)    0.265    0.023   11.515    0.000
##     a2         (g)    0.490    0.033   14.999    0.000
##   n2 ~                                                
##     xi2        (b)    0.265    0.023   11.515    0.000
##     a2         (g)    0.490    0.033   14.999    0.000
##   n3 ~                                                
##     xi3        (b)    0.265    0.023   11.515    0.000
##     a2         (g)    0.490    0.033   14.999    0.000
##   n4 ~                                                
##     xi4        (b)    0.265    0.023   11.515    0.000
##     a2         (g)    0.490    0.033   14.999    0.000
##   n5 ~                                                
##     xi5        (b)    0.265    0.023   11.515    0.000
##     a2         (g)    0.490    0.033   14.999    0.000
...
\end{verbatim}

\doublespacing

From this we can see that such a hybrid model is does a good job of
estimating the coefficients of interest, with \(\hat{\beta} =\) 0.265
(0.023) and \(\hat{\gamma} =\) 0.49 (0.033).

It is important, however, to realize that the unbiasedness of
\(\hat{\gamma}\) in this model is dependent on the assumption that
\(\mathop{\mathrm{\mathbb{E}}}[\zeta | \bm{\xi_{t}}, \alpha_{2}] = 0\).
In other words, the idiosyncratic error is mean independent of
\(\bm{\xi_{t}} = (\xi_{1}, \xi_{2}, ..., \xi_{T})\) as well as
\(\alpha_{2}\). The first part is easier to accept because we are
controlling for all potential time-invariant confounders that could
induce a relationship between the independent variable and the error.
The unbiasedness of \(\hat{\gamma}\), on the other hand rests on the
assumption that the time-invariant predictor is independent of the
error. If \(\alpha_{2}\) represented the respondent's intelligence and
\(\eta_{t}\), the dependent variable, represented the respondent's
income, for example, then \(\hat{\gamma}\) would be biased if both were
dependent on a third time-invariant variable, say level of schooling, if
it is not controlled for. For this reason, we need to treat the
regression on a time-invariant predictor like any other regular
multivariate regression model and look to include all plausible
potential confounders as controls in the model, or turn to other
methods, e.g., instrumental variables.

\bibliographystyle{tfcad}
\bibliography{references2.bib}

\end{document}